### 2015-11-16 latex/derivation.tex
Finish-ish velocities. They are probably mostly right. If wrong, we can come back later and fix them without wasting effort, since the actual values are now defined as coefficients A, B, C, etc...

\documentclass[11pt, landscape]{article}

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{fullpage}
\usepackage{verbatim}

\title{Brownian Dynein Model}

\newcommand{\mn}{\scalebox{0.7}[1.0]{-}}

\begin{document}

\maketitle

\section{Cartesian from Angular Coordinates}
This section lays out the definitions of the coordinates of the Dynein model. \\
\begin{align}
  X_{bb} &= 0 \\
  X_{bm} &= X_{bb}+L_{s}\cos(\Theta_{bb}) \\
  X_{t}  &= X_{bm}+L_{t}\cos(\Theta_{bm}) \\
  X_{fm} &= X_{t} - L_{t}\cos(\Theta_{fm}) \\
  X_{fb} &= X_{fm} - L_{s}\cos(\Theta_{fb})
\end{align}

\begin{align}
  Y_{bb} &= 0 \\
  Y_{bm} &= Y_{bb}+L_{s}\sin(\Theta_{bb}) \\
  Y_{t}  &= Y_{bm}+L_{t}\sin(\Theta_{bm}) \\
  Y_{fm} &= Y_{t} - L_{t}\sin(\Theta_{fm}) \\
  Y_{fb} &= Y_{fm} - L_{s}\sin(\Theta_{fb})
\end{align}

\section{Cartesian Velocities}
This section shows the velocity definitions. \\
\begin{align}
  \dot{X}_{bb} &= 0 \\
  \dot{X}_{bm} &= \dot{X}_{bb} - L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} \\
  \dot{X}_{t } &= \dot{X}_{bm} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} \\
  \dot{X}_{fm} &= \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} \\
  \dot{X}_{fb} &= \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}
\end{align}

\begin{align}                                                                          
  \dot{Y}_{bb} &= 0 \\                                                        
  \dot{Y}_{bm} &= \dot{Y}_{bb} + L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} \\
  \dot{Y}_{t}  &= \dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} \\
  \dot{Y}_{fm} &= \dot{Y}_{t } - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} \\
  \dot{Y}_{fb} &= \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}
\end{align}

\section{Cartesian Velocities (expanded)}
This section expands the above definitions. \\
\begin{align}
  \dot{X}_{bb} &= 0 \\
  \dot{X}_{bm} &= - L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} \\
  \dot{X}_{t } &= - L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} \\
  \dot{X}_{fm} &= - L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} \\
  \dot{X}_{fb} &= - L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}
\end{align}

\begin{align}                                                                               
  \dot{Y}_{bb} &= 0 \\                                                               
  \dot{Y}_{bm} &= L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} \\
  \dot{Y}_{t}  &= L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} \\
  \dot{Y}_{fm} &= L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} \\
  \dot{Y}_{fb} &= L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}
\end{align}

\section{Brownian Equations}
This section shows the velocity definitions using the Brownian dynamics equation. \\
\begin{align}  
  \dot{X}_{bb} &= 0 \\
  \dot{X}_{bm} &= \frac{1}{\gamma} \Big(F_{xml} + \mn \lambda_{bs}(X_{bm} - X_{bb}) + \lambda_{bt}(X_{t } - X_{bm}) \Big) + R_{xml} \\
  \dot{X}_{t } &= \frac{1}{\gamma} \Big(F_{xt } + \mn \lambda_{bt}(X_{t } - X_{bm}) + \lambda_{ft}(X_{fm} - X_{t }) \Big) + R_{xt } \\
  \dot{X}_{fm} &= \frac{1}{\gamma} \Big(F_{xmr} + \mn \lambda_{ft}(X_{fm} - X_{t }) + \lambda_{fs}(X_{fb} - X_{fm}) \Big) + R_{xmr} \\
  \dot{X}_{fb} &= \frac{1}{\gamma} \Big(F_{xbr} + \mn \lambda_{fs}(X_{fb} - X_{fm})                                 \Big) + R_{xbr}
\end{align}

\begin{align}  
  \dot{Y}_{bb} &= 0 \\
  \dot{Y}_{bm} &= \frac{1}{\gamma} \Big(F_{yml} + \mn \lambda_{bs}(Y_{bm} - Y_{bb}) + \lambda_{bt}(Y_{t } - Y_{bm}) \Big) + R_{yml} \\
  \dot{Y}_{t}  &= \frac{1}{\gamma} \Big(F_{yt } + \mn \lambda_{bt}(Y_{t } - Y_{bm}) + \lambda_{ft}(Y_{fm} - Y_{t }) \Big) + R_{yt } \\
  \dot{Y}_{fm} &= \frac{1}{\gamma} \Big(F_{ymr} + \mn \lambda_{ft}(Y_{fm} - Y_{t }) + \lambda_{fs}(Y_{fb} - Y_{fm}) \Big) + R_{ymr} \\
  \dot{Y}_{fb} &= \frac{1}{\gamma} \Big(F_{ybr} + \mn \lambda_{fs}(Y_{fb} - Y_{fm})                                 \Big) + R_{ybr}
\end{align}

%Unnecessary section commented out
\begin{comment}
\section{Expanded Brownian Equations}

\begin{align}
  &L_{s}\sin(\Theta_{bb})  \dot{\Theta}_{bb} = \frac{1}{\gamma}F_{xml} + \mn\frac{1}{\gamma}\lambda_{bs}(X_{bm} - X_{bb}) + \frac{1}{\gamma}\lambda_{bt}(X_{t } - X_{bm}) + R_{xml} \\
  &L_{s}\sin(\Theta_{bb})  \dot{\Theta}_{bb} + L_{t}\sin(\Theta_{bm})      \dot{\Theta}_{bm} = \frac{1}{\gamma}F_{xt } + \mn\frac{1}{\gamma}\lambda_{bt}(X_{t } - X_{bm}) + \frac{1}{\gamma}\lambda_{ft}(X_{fm} - X_{t }) + R_{xt } \\
  &L_{s}\sin(\Theta_{bb})  \dot{\Theta}_{bb} + L_{t}\sin(\Theta_{bm})      \dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm}-\pi)  \dot{\Theta}_{fm} = \frac{1}{\gamma}F_{xmr} + \mn\frac{1}{\gamma}\lambda_{ft}(X_{fm} - X_{t }) + \frac{1}{\gamma}\lambda_{fs}(X_{fb} - X_{fm}) + R_{xmr} \\
  &L_{s}\sin(\Theta_{bb})  \dot{\Theta}_{bb} + L_{t}\sin(\Theta_{bm})      \dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm}-\pi)  \dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb}-\pi)  \dot{\Theta}_{fb} = \frac{1}{\gamma}F_{xbr} + \mn\frac{1}{\gamma}\lambda_{fs}(X_{fb} - X_{fm}) + R_{xbr}
\end{align}

\begin{align}
  &\mn L_{s}\cos(\Theta_{bb})  \dot{\Theta}_{bb} = \frac{1}{\gamma}F_{yml} + \mn \frac{1}{\gamma}\lambda_{bs}(Y_{bm} - Y_{bb}) + \frac{1}{\gamma}\lambda_{bt}(Y_{t } - Y_{bm}) + R_{yml} \\
  &\mn L_{s}\cos(\Theta_{bb})  \dot{\Theta}_{bb} + \mn L_{t}\cos(\Theta_{bm})  \dot{\Theta}_{bm}  = \frac{1}{\gamma}F_{yt } + \mn \frac{1}{\gamma}\lambda_{bt}(Y_{t } - Y_{bm}) + \frac{1}{\gamma}\lambda_{ft}(Y_{fm} - Y_{t }) + R_{yt } \\
  &\mn L_{s}\cos(\Theta_{bb})  \dot{\Theta}_{bb} + \mn L_{t}\cos(\Theta_{bm})  \dot{\Theta}_{bm} + \mn L_{t}\cos(\Theta_{fm}-\pi)  \dot{\Theta}_{fm} = \frac{1}{\gamma}F_{ymr} + \mn \frac{1}{\gamma}\lambda_{ft}(Y_{fm} - Y_{t }) + \frac{1}{\gamma}\lambda_{fs}(Y_{fb} - Y_{fm}) + R_{ymr} \\
  &\mn L_{s}\cos(\Theta_{bb})  \dot{\Theta}_{bb} + \mn L_{t}\cos(\Theta_{bm})  \dot{\Theta}_{bm} + \mn L_{t}\cos(\Theta_{fm}-\pi)  \dot{\Theta}_{fm} + \mn L_{s}\cos(\Theta_{fb}-\pi)  \dot{\Theta}_{fb} = \frac{1}{\gamma}F_{ybr} + \mn \frac{1}{\gamma}\lambda_{fs}(Y_{fb} - Y_{fm}) + R_{ybr}
\end{align}

\section{System after Subtracting out Angular Velocity Terms}
This is the system after subtracting equations from each other to eliminate some angular velocity terms.\\

\begin{align}
  L_{s}\sin(\Theta_{bb})      \dot{\Theta}_{bb} &= \frac{1}{\gamma}F_{xml} - \frac{1}{\gamma}\lambda_{bs}(X_{bm} - X_{bb}) + \frac{1}{\gamma}\lambda_{bt}(X_{t } - X_{bm}) + R_{xml} \\
  L_{t}\sin(\Theta_{bm})      \dot{\Theta}_{bm} &= \frac{1}{\gamma}F_{xt } - \frac{2}{\gamma}\lambda_{bt}(X_{t } - X_{bm}) + \frac{1}{\gamma}\lambda_{ft}(X_{fm} - X_{t }) + R_{xt } - \frac{1}{\gamma}F_{xml} - \mn\frac{1}{\gamma}\lambda_{bs}(X_{bm} - X_{bb}) - R_{xml} \\
  L_{t}\sin(\Theta_{fm}-\pi)  \dot{\Theta}_{fm} &= \frac{1}{\gamma}F_{xmr} - \frac{2}{\gamma}\lambda_{ft}(X_{fm} - X_{t }) + \frac{1}{\gamma}\lambda_{fs}(X_{fb} - X_{fm}) + R_{xmr} - \frac{1}{\gamma}F_{xt } - \mn\frac{1}{\gamma}\lambda_{bt}(X_{t } - X_{bm}) - R_{xt } \\
  L_{s}\sin(\Theta_{fb}-\pi)  \dot{\Theta}_{fb} &= \frac{1}{\gamma}F_{xbr} - \frac{2}{\gamma}\lambda_{fs}(X_{fb} - X_{fm}) + R_{xbr} - \frac{1}{\gamma}F_{xmr} - \mn\frac{1}{\gamma}\lambda_{ft}(X_{fm} - X_{t }) - R_{xmr} 
\end{align}

\begin{align}
  \mn L_{s}\cos(\Theta_{bb})      \dot{\Theta}_{bb} &= \frac{1}{\gamma}F_{yml} - \frac{1}{\gamma}\lambda_{bs}(Y_{bm} - Y_{bb}) + \frac{1}{\gamma}\lambda_{bt}(Y_{t } - Y_{bm}) + R_{yml} \\
  \mn L_{t}\cos(\Theta_{bm})      \dot{\Theta}_{bm} &= \frac{1}{\gamma}F_{yt } - \frac{2}{\gamma}\lambda_{bt}(Y_{t } - Y_{bm}) + \frac{1}{\gamma}\lambda_{ft}(Y_{fm} - Y_{t }) + R_{yt } - \frac{1}{\gamma}F_{yml} - \mn \frac{1}{\gamma}\lambda_{bs}(Y_{bm} - Y_{bb}) - R_{yml}\\
  \mn L_{t}\cos(\Theta_{fm}-\pi)  \dot{\Theta}_{fm} &= \frac{1}{\gamma}F_{ymr} - \frac{2}{\gamma}\lambda_{ft}(Y_{fm} - Y_{t }) + \frac{1}{\gamma}\lambda_{fs}(Y_{fb} - Y_{fm}) + R_{ymr} - \frac{1}{\gamma}F_{yt } - \mn \frac{1}{\gamma}\lambda_{bt}(Y_{t } - Y_{bm}) - R_{yt } \\
  \mn L_{s}\cos(\Theta_{fb}-\pi)  \dot{\Theta}_{fb} &= \frac{1}{\gamma}F_{ybr} - \frac{2}{\gamma}\lambda_{fs}(Y_{fb} - Y_{fm}) + R_{ybr} - \frac{1}{\gamma}F_{ymr} - \mn \frac{1}{\gamma}\lambda_{ft}(Y_{fm} - Y_{t }) - R_{ymr} 
\end{align}
\end{comment}

\section{System of Equations}
This section creates a system of equations by equating the Cartesian and Brownian velocity definitions from above. \\

\begin{align}  
  \dot{X}_{bm} &= \dot{X}_{bb} + \mn L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} = \frac{1}{\gamma}F_{xml} + \mn\frac{1}{\gamma}\lambda_{bs}(X_{bm} - X_{bb}) + \frac{1}{\gamma}\lambda_{bt}(X_{t } - X_{bm}) + R_{xml} \\
  \dot{X}_{t}  &= \dot{X}_{bm} + \mn L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} = \frac{1}{\gamma}F_{xt } + \mn\frac{1}{\gamma}\lambda_{bt}(X_{t } - X_{bm}) + \frac{1}{\gamma}\lambda_{ft}(X_{fm} - X_{t }) + R_{xt } \\
  \dot{X}_{fm} &= \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} = \frac{1}{\gamma}F_{xmr} + \mn\frac{1}{\gamma}\lambda_{ft}(X_{fm} - X_{t }) + \frac{1}{\gamma}\lambda_{fs}(X_{fb} - X_{fm}) + R_{xmr} \\
  \dot{X}_{fb} &= \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} = \frac{1}{\gamma}F_{xbr} + \mn\frac{1}{\gamma}\lambda_{fs}(X_{fb} - X_{fm}) + R_{xbr}
\end{align}

\begin{align}  
  \dot{Y}_{bm} &= \dot{Y}_{bb} + L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} = \frac{1}{\gamma}F_{yml} + \mn\frac{1}{\gamma}\lambda_{bs}(Y_{bm} - Y_{bb}) + \frac{1}{\gamma}\lambda_{bt}(Y_{t } - Y_{bm}) + R_{yml} \\
  \dot{Y}_{t}  &= \dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} = \frac{1}{\gamma}F_{yt } + \mn\frac{1}{\gamma}\lambda_{bt}(Y_{t } - Y_{bm}) + \frac{1}{\gamma}\lambda_{ft}(Y_{fm} - Y_{t }) + R_{yt } \\
  \dot{Y}_{fm} &= \dot{Y}_{t } + \mn L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} = \frac{1}{\gamma}F_{ymr} + \mn\frac{1}{\gamma}\lambda_{ft}(Y_{fm} - Y_{t }) + \frac{1}{\gamma}\lambda_{fs}(Y_{fb} - Y_{fm}) + R_{ymr} \\
  \dot{Y}_{fb} &= \dot{Y}_{fm} + \mn L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} = \frac{1}{\gamma}F_{ybr} + \mn\frac{1}{\gamma}\lambda_{fs}(Y_{fb} - Y_{fm}) + R_{ybr}
\end{align}

\section{System after Addition}
By adding equations from our system together we can isolate the $\lambda$ values. This system has been verified via CAS.\\

\begin{multline}
\mn\frac{1}{\gamma}\lambda_{bs}(X_{bm} - X_{bb}) =
\dot{X}_{bb} + \mn L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - R_{xml} - \frac{1}{\gamma}F_{xml} + \dot{X}_{bm} + \mn L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} - R_{xt } \\
- \frac{1}{\gamma}F_{xt } + \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} - R_{xmr}
-\frac{1}{\gamma}F_{xmr} + \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr}
\label{eq:Ax}
\end{multline}%Fix'd

\begin{multline}
\mn\frac{1}{\gamma}\lambda_{bt}(X_{t } - X_{bm}) =
\dot{X}_{bm} + \mn L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} - R_{xt } - \frac{1}{\gamma}F_{xt } + \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} \\
- R_{xmr} - \frac{1}{\gamma}F_{xmr} + \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr}
\label{eq:Bx}
\end{multline}%Fix'd

\begin{equation}
\mn\frac{1}{\gamma}\lambda_{ft}(X_{fm} - X_{t }) =
\dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} - R_{xmr} - \frac{1}{\gamma}F_{xmr}
+ \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr}
\label{eq:Cx}
\end{equation}%Fix'd

\begin{equation}
\mn\frac{1}{\gamma}\lambda_{fs}(X_{fb} - X_{fm}) = 
\dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr}
\label{eq:Dx}
\end{equation}%Fix'd

\begin{multline}
\mn\frac{1}{\gamma}\lambda_{bs}(Y_{bm} - Y_{bb}) =
\dot{Y}_{bb} + L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} - R_{yml} - \frac{1}{\gamma}F_{yml} + \dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} \\
- R_{yt} - \frac{1}{\gamma}F_{yt } + \dot{Y}_{t } + \mn L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - R_{ymr} - \frac{1}{\gamma}F_{ymr} + \dot{Y}_{fm}
+ \mn L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}
\label{eq:Ay}
\end{multline}%Fix'd

\begin{multline}
\mn\frac{1}{\gamma}\lambda_{bt}(Y_{t } - Y_{bm}) =
\dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - R_{yt} - \frac{1}{\gamma}F_{yt } + \dot{Y}_{t } + \mn L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} \\
- R_{ymr} - \frac{1}{\gamma}F_{ymr} + \dot{Y}_{fm} + \mn L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}
\label{eq:By}
\end{multline}%Fix'd

\begin{equation}
\mn\frac{1}{\gamma}\lambda_{ft}(Y_{fm} - Y_{t }) =
\dot{Y}_{t } + \mn L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - R_{ymr} - \frac{1}{\gamma}F_{ymr} + \dot{Y}_{fm} + \mn L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}
- R_{ybr} - \frac{1}{\gamma}F_{ybr}
\label{eq:Cy}
\end{equation}%Fix'd

\begin{equation}
\mn\frac{1}{\gamma}\lambda_{fs}(Y_{fb} - Y_{fm}) = \dot{Y}_{fm} + \mn L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}
\label{eq:Dy}
\end{equation}%Fix'd

\section{Creating Angular Velocity Matrix}

Our goal is to take the above system of equations and put it in matrix form:

\begin{align}
  A_1\dot{\Theta}_{bb} + A_2 \dot{\Theta}_{bm} + A_3 \dot{\Theta}_{fm} + A_4 \dot{\Theta}_{fb} &= X_1 \\
  B_1\dot{\Theta}_{bb} + B_2 \dot{\Theta}_{bm} + B_3 \dot{\Theta}_{fm} + B_4 \dot{\Theta}_{fb} &= X_2 \\
  C_1\dot{\Theta}_{bb} + C_2 \dot{\Theta}_{bm} + C_3 \dot{\Theta}_{fm} + C_4 \dot{\Theta}_{fb} &= X_3 \\
  D_1\dot{\Theta}_{bb} + D_2 \dot{\Theta}_{bm} + D_3 \dot{\Theta}_{fm} + D_4 \dot{\Theta}_{fb} &= X_4 \\
\end{align}

This will allow us to perform a matrix solution for the angular velocities. $X_1$, $X_2$, $X_3$ and $X_4$ contain only force terms and no unknown $\lambda$s. The above system of equations has four lambda terms: $\lambda_{bs}, \lambda_{bt}, \lambda_{ft}$ and $\lambda_{bs}$. Each term is present in two equations. We will use this redundancy to eliminate the lambda terms and create the above matrix.

\subsection{Solving for $A_1$, $A_2$, $A_3$, $A_4$ and $X_1$}
In this section we take Equations (\ref{eq:Ax}) and (\ref{eq:Ay}), equate them and manipulate them to be in form $A_1\dot{\Theta}_{bb} + A_2 \dot{\Theta}_{bm} + A_3 \dot{\Theta}_{fm} + A_4 \dot{\Theta}_{fb} = X_1$\\

We start by dividing out the position terms in Equations (\ref{eq:Ax}) and (\ref{eq:Ay}) to make their LHSs identical.

\begin{multline}
\mn\frac{1}{\gamma}\lambda_{bs} =
\Big(\dot{X}_{bb} + \mn L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - R_{xml} - \frac{1}{\gamma}F_{xml} + \dot{X}_{bm} + \mn L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} - R_{xt } \\
- \frac{1}{\gamma}F_{xt } + \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} - R_{xmr}
-\frac{1}{\gamma}F_{xmr} + \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr}\Big)\frac{1}{(X_{bm} - X_{bb})}
\end{multline}

\begin{multline}
\mn\frac{1}{\gamma}\lambda_{bs} =
\Big( \dot{Y}_{bb} + L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} - R_{yml} - \frac{1}{\gamma}F_{yml} + \dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} \\
- R_{yt} - \frac{1}{\gamma}F_{yt } + \dot{Y}_{t } + \mn L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - R_{ymr} - \frac{1}{\gamma}F_{ymr} + \dot{Y}_{fm}
+ \mn L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\Big)\frac{1}{(Y_{bm} - Y_{bb})}
\end{multline}

We then equate them.

\begin{multline}
\Big(\dot{X}_{bb} + \mn L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - R_{xml} - \frac{1}{\gamma}F_{xml} + \dot{X}_{bm} + \mn L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} - R_{xt } \\
- \frac{1}{\gamma}F_{xt } + \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} - R_{xmr}
-\frac{1}{\gamma}F_{xmr} + \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr}\Big)(Y_{bm} - Y_{bb}) = \\
\Big( \dot{Y}_{bb} + L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} - R_{yml} - \frac{1}{\gamma}F_{yml} + \dot{Y}_{bm} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm}
- R_{yt} - \frac{1}{\gamma}F_{yt } + \dot{Y}_{t } + \mn L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - R_{ymr} - \frac{1}{\gamma}F_{ymr} + \dot{Y}_{fm} \\
+ \mn L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\Big)(X_{bm} - X_{bb})
\end{multline}

Then we group like terms for clarity.

\begin{multline}
\Big(\dot{X}_{bb} + \dot{X}_{bm} + \dot{X}_{t } + \dot{X}_{fm} 
- L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} \\
- \frac{1}{\gamma}F_{xml} - \frac{1}{\gamma}F_{xt } -\frac{1}{\gamma}F_{xmr} - \frac{1}{\gamma}F_{xbr}
- R_{xml} - R_{xt } - R_{xmr} - R_{xbr} \Big)(L_s\sin(\Theta_{bb})) = \\
\Big(\dot{Y}_{bb} + \dot{Y}_{bm} + \dot{Y}_{t } + \dot{Y}_{fm}
+ L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} \\
- \frac{1}{\gamma}F_{yml} - \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}
- R_{yml} - R_{yt} - R_{ymr} - R_{ybr}\Big)(L_s\cos(\Theta_{bb}))
\end{multline}

We then expand all the $\dot{X}$ terms using their definitions from the Cartesian Velocities section above.

\begin{multline}
\Big(
- L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb}
- L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm}
- L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm}\\
- L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}
- \frac{1}{\gamma}F_{xml} - \frac{1}{\gamma}F_{xt } -\frac{1}{\gamma}F_{xmr} - \frac{1}{\gamma}F_{xbr}
- R_{xml} - R_{xt } - R_{xmr} - R_{xbr} \Big)(L_s\sin(\Theta_{bb})) = \\
\Big(
+ L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb}
+ L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm}
+ L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm}\\
+ L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}
- \frac{1}{\gamma}F_{yml} - \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}
- R_{yml} - R_{yt} - R_{ymr} - R_{ybr}\Big)(L_s\cos(\Theta_{bb}))
\end{multline}

Then we group together the like terms from the expansion.

\begin{multline}
\Big(- 4L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - 3L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + 2L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}
- \frac{1}{\gamma}F_{xml} - \frac{1}{\gamma}F_{xt } -\frac{1}{\gamma}F_{xmr} - \frac{1}{\gamma}F_{xbr}
- R_{xml} - R_{xt } - R_{xmr} - R_{xbr} \Big)(L_s\sin(\Theta_{bb})) = \\
\Big(4L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + 3L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - 2L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}
- \frac{1}{\gamma}F_{yml} - \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}
- R_{yml} - R_{yt} - R_{ymr} - R_{ybr}\Big)(L_s\cos(\Theta_{bb}))
\end{multline}

We then bring all the angular velocity terms to one side and all the force terms to the other side.

\begin{multline}
\Big(- 4L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - 3L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm}
+ 2L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}\Big)(L_s\sin(\Theta_{bb}))\\
\Big(-4L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} - 3L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm}
+ 2L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\Big)(L_s\cos(\Theta_{bb}))= \\
\Big(- \frac{1}{\gamma}F_{yml} - \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yml} - R_{yt} - R_{ymr} - R_{ybr}\Big)(L_s\cos(\Theta_{bb}))
+ \Big( \frac{1}{\gamma}F_{xml} + \frac{1}{\gamma}F_{xt } + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr} + R_{xml} + R_{xt } + R_{xmr} + R_{xbr} \Big)(L_s\sin(\Theta_{bb}))
\end{multline}

We then distribute the $L_s\sin(\Theta_{bb})$ and $L_s\cos(\Theta_{bb})$ terms into the expression.

\begin{multline}
\Big(- 4L_{s}\sin(\Theta_{bb})\sin(\Theta_{bb})\dot{\Theta}_{bb} - 3L_{t}\sin(\Theta_{bm})\sin(\Theta_{bb})\dot{\Theta}_{bm}
+ 2L_{t}\sin(\Theta_{fm})\sin(\Theta_{bb})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\sin(\Theta_{bb})\dot{\Theta}_{fb}\Big)\\
\Big(-4L_{s}\cos(\Theta_{bb})\cos(\Theta_{bb})\dot{\Theta}_{bb} - 3L_{t}\cos(\Theta_{bm})\cos(\Theta_{bb})\dot{\Theta}_{bm}
+ 2L_{t}\cos(\Theta_{fm})\cos(\Theta_{bb})\dot{\Theta}_{fm} + L_{s}\cos(\Theta_{fb})\cos(\Theta_{bb})\dot{\Theta}_{fb}\Big)= \\
\Big(- \frac{1}{\gamma}F_{yml} - \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yml} - R_{yt} - R_{ymr} - R_{ybr}\Big)(\cos(\Theta_{bb}))
+ \Big( \frac{1}{\gamma}F_{xml} + \frac{1}{\gamma}F_{xt } + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr} + R_{xml} + R_{xt } + R_{xmr} + R_{xbr} \Big)(\sin(\Theta_{bb}))
\end{multline}

Finally, we group all terms by their angular velocity.

\begin{multline}
-4L_{s}\Big(\sin(\Theta_{bb})\sin(\Theta_{bb}) + \cos(\Theta_{bb})\cos(\Theta_{bb})\Big)\dot{\Theta}_{bb}
-3L_{t}\Big(\sin(\Theta_{bm})\sin(\Theta_{bb}) + \cos(\Theta_{bm})\cos(\Theta_{bb})\Big)\dot{\Theta}_{bm}\\
+2L_{t}\Big(\sin(\Theta_{fm})\sin(\Theta_{bb}) + \cos(\Theta_{fm})\cos(\Theta_{bb})\Big)\dot{\Theta}_{fm}
+ L_{s}\Big(\sin(\Theta_{fb})\sin(\Theta_{bb}) + \cos(\Theta_{fb})\cos(\Theta_{bb})\Big)\dot{\Theta}_{fb}= \\
\Big(- \frac{1}{\gamma}F_{yml} - \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yml} - R_{yt} - R_{ymr} - R_{ybr}\Big)(\cos(\Theta_{bb}))
+ \Big( \frac{1}{\gamma}F_{xml} + \frac{1}{\gamma}F_{xt } + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr} + R_{xml} + R_{xt } + R_{xmr} + R_{xbr} \Big)(\sin(\Theta_{bb}))
\end{multline}

This is the form we want. We now know the value of $A_1, A_2, A_3, A_4$ and $X_1$.

\begin{align}
  A_1 &= -4L_{s}\\
  A_2 &= -3L_{t}\Big(\sin(\Theta_{bm})\sin(\Theta_{bb}) + \cos(\Theta_{bm})\cos(\Theta_{bb})\Big)\\
  A_3 &= 2L_{t}\Big(\sin(\Theta_{fm})\sin(\Theta_{bb}) + \cos(\Theta_{fm})\cos(\Theta_{bb})\Big)\\
  A_4 &= L_{s}\Big(\sin(\Theta_{fb})\sin(\Theta_{bb}) + \cos(\Theta_{fb})\cos(\Theta_{bb})\Big)\\
  X_1 &= \Big(- \frac{1}{\gamma}F_{yml} - \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yml} - R_{yt} - R_{ymr} - R_{ybr}\Big)(\cos(\Theta_{bb}))+\\
      &+ \Big( \frac{1}{\gamma}F_{xml} + \frac{1}{\gamma}F_{xt } + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr} + R_{xml} + R_{xt } + R_{xmr} + R_{xbr} \Big)(\sin(\Theta_{bb}))
\end{align}

\subsection{Solving for $B_1$, $B_2$, $B_3$, $B_4$ and $X_2$}
Here we take Equations (\ref{eq:Bx}) and (\ref{eq:By}), equate them and manipulate them to be in form $B_1\dot{\Theta}_{bb} + B_2 \dot{\Theta}_{bm} + B_3 \dot{\Theta}_{fm} + B_4 \dot{\Theta}_{fb} = X_2$

\begin{multline}
-\frac{1}{\gamma}\lambda_{bt} =
\Big(
\dot{X}_{bm} + \dot{X}_{t } + \dot{X}_{fm}
- L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} 
- \frac{1}{\gamma}F_{xt } - \frac{1}{\gamma}F_{xmr} - \frac{1}{\gamma}F_{xbr}
- R_{xt } - R_{xmr} - R_{xbr}\Big) \frac{1}{(X_{t } - X_{bm})}
\end{multline}

\begin{multline}
-\frac{1}{\gamma}\lambda_{bt} =
\Big(
\dot{Y}_{bm} + \dot{Y}_{t } + \dot{Y}_{fm} 
+ L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} 
- \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}
- R_{yt} - R_{ymr} - R_{ybr}\Big) \frac{1}{(Y_{t } - Y_{bm})}
\end{multline}

\begin{multline}
\Big(
\dot{X}_{bm} + \dot{X}_{t } + \dot{X}_{fm}
- L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} 
- \frac{1}{\gamma}F_{xt } - \frac{1}{\gamma}F_{xmr} - \frac{1}{\gamma}F_{xbr}
- R_{xt } - R_{xmr} - R_{xbr}\Big)(Y_{t } - Y_{bm}) =\\
\Big(
\dot{Y}_{bm} + \dot{Y}_{t } + \dot{Y}_{fm} 
+ L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} 
- \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}
- R_{yt} - R_{ymr} - R_{ybr}\Big)(X_{t } - X_{bm})
\end{multline}

\begin{multline}
\Big(
\dot{X}_{bm} + \dot{X}_{t } + \dot{X}_{fm}
- L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} 
- \frac{1}{\gamma}F_{xt } - \frac{1}{\gamma}F_{xmr} - \frac{1}{\gamma}F_{xbr}
- R_{xt } - R_{xmr} - R_{xbr}\Big)(L_t\sin(\Theta_{bm})) =\\
\Big(
\dot{Y}_{bm} + \dot{Y}_{t } + \dot{Y}_{fm} 
+ L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} 
- \frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}
- R_{yt} - R_{ymr} - R_{ybr}\Big)(L_t\cos(\Theta_{bm}))
\end{multline}

\begin{multline}
\Big(\dot{X}_{bm} + \dot{X}_{t } + \dot{X}_{fm}
- L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}\Big)(L_t\sin(\Theta_{bm}))\\
- \Big(\dot{Y}_{bm} + \dot{Y}_{t } + \dot{Y}_{fm} 
+ L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\Big)(L_t\cos(\Theta_{bm})) =\\
\Big(-\frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yt} - R_{ymr} - R_{ybr}\Big)(L_t\cos(\Theta_{bm}))
- \Big(-\frac{1}{\gamma}F_{xt } - \frac{1}{\gamma}F_{xmr} - \frac{1}{\gamma}F_{xbr} - R_{xt } - R_{xmr} - R_{xbr}\Big)(L_t\sin(\Theta_{bm}))
\end{multline}

\begin{multline}
\Big( -3L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - 3L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + 2L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm}
+ L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}\Big)(L_t\sin(\Theta_{bm}))\\
- \Big( 3L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + 3L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} -2L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm}
- L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\Big)(L_t\cos(\Theta_{bm})) =\\
\Big(-\frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yt} - R_{ymr} - R_{ybr}\Big)(L_t\cos(\Theta_{bm}))
- \Big(-\frac{1}{\gamma}F_{xt } - \frac{1}{\gamma}F_{xmr} - \frac{1}{\gamma}F_{xbr} - R_{xt } - R_{xmr} - R_{xbr}\Big)(L_t\sin(\Theta_{bm}))
\end{multline}

\begin{multline}
\Big( -3L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - 3L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + 2L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm}
+ L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}\Big)(L_t\sin(\Theta_{bm}))\\
+ \Big(-3L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} - 3L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} + 2L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm}
+ L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\Big)(L_t\cos(\Theta_{bm})) =\\
\Big(-\frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yt} - R_{ymr} - R_{ybr}\Big)(L_t\cos(\Theta_{bm}))
+ \Big(\frac{1}{\gamma}F_{xt } + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr} + R_{xt } + R_{xmr} + R_{xbr}\Big)(L_t\sin(\Theta_{bm}))
\end{multline}

\begin{multline}
\Big( -3L_{s}\sin(\Theta_{bb})L_t\sin(\Theta_{bm})\dot{\Theta}_{bb} - 3L_{t}\sin(\Theta_{bm})L_t\sin(\Theta_{bm})\dot{\Theta}_{bm}
+ 2L_{t}\sin(\Theta_{fm})L_t\sin(\Theta_{bm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})L_t\sin(\Theta_{bm})\dot{\Theta}_{fb}\Big)\\
+ \Big(-3L_{s}\cos(\Theta_{bb})L_t\cos(\Theta_{bm})\dot{\Theta}_{bb} - 3L_{t}\cos(\Theta_{bm})L_t\cos(\Theta_{bm})\dot{\Theta}_{bm}
+ 2L_{t}\cos(\Theta_{fm})L_t\cos(\Theta_{bm})\dot{\Theta}_{fm} + L_{s}\cos(\Theta_{fb})L_t\cos(\Theta_{bm})\dot{\Theta}_{fb}\Big) =\\
\Big(-\frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yt} - R_{ymr} - R_{ybr}\Big)(L_t\cos(\Theta_{bm}))
+ \Big(\frac{1}{\gamma}F_{xt } + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr} + R_{xt } + R_{xmr} + R_{xbr}\Big)(L_t\sin(\Theta_{bm}))
\end{multline}

\begin{multline}
-3L_{s}\Big(\sin(\Theta_{bb})\sin(\Theta_{bm}) + \cos(\Theta_{bb})\cos(\Theta_{bm})\Big)\dot{\Theta}_{bb}
-3L_{t}\Big(\sin(\Theta_{bm})\sin(\Theta_{bm}) + \cos(\Theta_{bm})\cos(\Theta_{bm})\Big)\dot{\Theta}_{bm}\\
+2L_{t}\Big(\sin(\Theta_{fm})\sin(\Theta_{bm}) + \cos(\Theta_{fm})\cos(\Theta_{bm})\Big)\dot{\Theta}_{fm}
+ L_{s}\Big(\sin(\Theta_{fb})\sin(\Theta_{bm}) + \cos(\Theta_{fb})\cos(\Theta_{bm})\Big)\dot{\Theta}_{fb} =\\
\Big(-\frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yt} - R_{ymr} - R_{ybr}\Big)(\cos(\Theta_{bm}))
+ \Big(\frac{1}{\gamma}F_{xt } + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr} + R_{xt } + R_{xmr} + R_{xbr}\Big)(\sin(\Theta_{bm}))
\end{multline}

\begin{align}
  B_1 &= -3L_{s}\Big(\sin(\Theta_{bb})\sin(\Theta_{bm}) + \cos(\Theta_{bb})\cos(\Theta_{bm})\Big)\\
  B_2 &= -3L_{t}\\
  B_3 &= +2L_{t}\Big(\sin(\Theta_{fm})\sin(\Theta_{bm}) + \cos(\Theta_{fm})\cos(\Theta_{bm})\Big)\\
  B_4 &= + L_{s}\Big(\sin(\Theta_{fb})\sin(\Theta_{bm}) + \cos(\Theta_{fb})\cos(\Theta_{bm})\Big)\\
  X_2 &= \Big(-\frac{1}{\gamma}F_{yt } - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr} - R_{yt} - R_{ymr} - R_{ybr}\Big)(\cos(\Theta_{bm}))\\
      &+ \Big(\frac{1}{\gamma}F_{xt } + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr} + R_{xt } + R_{xmr} + R_{xbr}\Big)(\sin(\Theta_{bm}))
\end{align}

\subsection{Solving for $C_1$, $C_2$, $C_3$, $C_4$ and $X_3$}
Here we take Equations (\ref{eq:Cx}) and (\ref{eq:Cy}), equate them and manipulate them to be in form $ C_1\dot{\Theta}_{bb} + C_2 \dot{\Theta}_{bm} + C_3 \dot{\Theta}_{fm} + C_4 \dot{\Theta}_{fb} = X_3$

\begin{equation}
-\frac{1}{\gamma}\lambda_{ft} = \Big( \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} - R_{xmr} - \frac{1}{\gamma}F_{xmr}
+ \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr}\Big) \frac{1}{(X_{fm} - X_{t })}
\end{equation}

\begin{equation}
-\frac{1}{\gamma}\lambda_{ft} = \Big( \dot{Y}_{t } - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - R_{ymr} - \frac{1}{\gamma}F_{ymr}
+ \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\Big) \frac{1}{(Y_{fm} - Y_{t })}
\end{equation}

\begin{multline}
\Big( \dot{X}_{t } + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} - R_{xmr} - \frac{1}{\gamma}F_{xmr}
+ \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr}\Big)(-L_{t}\sin(\Theta_{fm})) = \\
\Big( \dot{Y}_{t } - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - R_{ymr} - \frac{1}{\gamma}F_{ymr}
+ \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\Big)(-L_{t}\cos(\Theta_{fm}))
\end{multline}

\begin{multline}
\Big( \dot{X}_{t } + \dot{X}_{fm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} \Big)\sin(\Theta_{fm}) +
\Big(- R_{xmr} - \frac{1}{\gamma}F_{xmr} - R_{xbr} - \frac{1}{\gamma}F_{xbr}\Big)\sin(\Theta_{fm}) = \\
\Big(\dot{Y}_{t } + \dot{Y}_{fm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\Big)\cos(\Theta_{fm}) +
\Big(- R_{ymr} - \frac{1}{\gamma}F_{ymr} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\Big)\cos(\Theta_{fm})
\end{multline}

\begin{multline}
\Big( \dot{X}_{t } + \dot{X}_{fm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} \Big)\sin(\Theta_{fm})
+ \Big(-\dot{Y}_{t } - \dot{Y}_{fm} + L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\Big)\cos(\Theta_{fm}) = \\
\Big(- R_{ymr} - \frac{1}{\gamma}F_{ymr} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\Big)\cos(\Theta_{fm})
+ \Big(R_{xmr} + \frac{1}{\gamma}F_{xmr} + R_{xbr} + \frac{1}{\gamma}F_{xbr}\Big)\sin(\Theta_{fm})
\end{multline}

\begin{multline}
\Big(- L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + -L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} +
L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} +L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} \Big)\sin(\Theta_{fm}) \\
+ \Big(-(L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm}) -(L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} -
L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm}) + L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\Big)\cos(\Theta_{fm}) =\\
\Big(-R_{ymr} - \frac{1}{\gamma}F_{ymr} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\Big)\cos(\Theta_{fm}) + \Big(R_{xmr} + \frac{1}{\gamma}F_{xmr} + R_{xbr}
+ \frac{1}{\gamma}F_{xbr}\Big)\sin(\Theta_{fm})
\end{multline}

\begin{multline}
\Big(-2L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} - 2L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} + 2L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm}
+ L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} \Big)\sin(\Theta_{fm}) \\
+ \Big(-2L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} - 2L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} + 2L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm}
+ L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\Big)\cos(\Theta_{fm}) =\\
\Big(-R_{ymr} -R_{ybr} - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}\Big)\cos(\Theta_{fm}) + \Big(R_{xmr} + R_{xbr} + \frac{1}{\gamma}F_{xmr}
+ \frac{1}{\gamma}F_{xbr}\Big)\sin(\Theta_{fm})
\end{multline}

\begin{multline}
\Big(-2L_{s}\sin(\Theta_{bb})\sin(\Theta_{fm})\dot{\Theta}_{bb} - 2L_{t}\sin(\Theta_{bm})\sin(\Theta_{fm})\dot{\Theta}_{bm}
+ 2L_{t}\sin(\Theta_{fm})\sin(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\sin(\Theta_{fb})\sin(\Theta_{fm})\dot{\Theta}_{fb} \Big)\\
+ \Big(-2L_{s}\cos(\Theta_{bb})\cos(\Theta_{fm})\dot{\Theta}_{bb} - 2L_{t}\cos(\Theta_{bm})\cos(\Theta_{fm})\dot{\Theta}_{bm}
+ 2L_{t}\cos(\Theta_{fm})\cos(\Theta_{fm})\dot{\Theta}_{fm} + L_{s}\cos(\Theta_{fb})\cos(\Theta_{fm})\dot{\Theta}_{fb}\Big) =\\
\Big(-R_{ymr} -R_{ybr} - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}\Big)\cos(\Theta_{fm}) + \Big(R_{xmr} + R_{xbr} + \frac{1}{\gamma}F_{xmr}
+ \frac{1}{\gamma}F_{xbr}\Big)\sin(\Theta_{fm})
\end{multline}

\begin{multline}
- 2L_{s}\Big(\sin(\Theta_{bb})\sin(\Theta_{fm}) + \cos(\Theta_{bb})\cos(\Theta_{fm})\Big)\dot{\Theta}_{bb}
- 2L_{t}\Big(\sin(\Theta_{bm})\sin(\Theta_{fm}) + \cos(\Theta_{bm})\cos(\Theta_{fm})\Big)\dot{\Theta}_{bm}\\
+ 2L_{t}\Big(\sin^2(\Theta_{fm}) + \cos^2(\Theta_{fm})\Big)\dot{\Theta}_{fm}
+ L_{s}\Big(\sin(\Theta_{fb})\sin(\Theta_{fm}) + \cos(\Theta_{fb})\cos(\Theta_{fm})\Big)\dot{\Theta}_{fb} =\\
\Big(-R_{ymr} -R_{ybr} - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}\Big)\cos(\Theta_{fm}) + \Big(R_{xmr} + R_{xbr} + \frac{1}{\gamma}F_{xmr}
+ \frac{1}{\gamma}F_{xbr}\Big)\sin(\Theta_{fm})
\end{multline}

\begin{align}
  C_1 &= - 2L_{s}\Big(\sin(\Theta_{bb})\sin(\Theta_{fm}) + \cos(\Theta_{bb})\cos(\Theta_{fm})\Big) \\
  C_2 &= - 2L_{t}\Big(\sin(\Theta_{bm})\sin(\Theta_{fm}) + \cos(\Theta_{bm})\cos(\Theta_{fm})\Big) \\
  C_3 &= 2L_{t} \\
  C_4 &= L_{s}\Big(\sin(\Theta_{fb})\sin(\Theta_{fm}) + \cos(\Theta_{fb})\cos(\Theta_{fm})\Big) \\
  X_3 &= \Big(-R_{ymr} -R_{ybr} - \frac{1}{\gamma}F_{ymr} - \frac{1}{\gamma}F_{ybr}\Big)\cos(\Theta_{fm})\\
      &+ \Big(R_{xmr} + R_{xbr} + \frac{1}{\gamma}F_{xmr} + \frac{1}{\gamma}F_{xbr}\Big)\sin(\Theta_{fm})
\end{align}

\subsection{Solving for $D_1$, $D_2$, $D_3$, $D_4$ and $X_4$}
Here we take Equations (\ref{eq:Dx}) and (\ref{eq:Dy}), equate them and manipulate them to be in form $D_1\dot{\Theta}_{bb} + D_2 \dot{\Theta}_{bm} + D_3 \dot{\Theta}_{fm} + D_4 \dot{\Theta}_{fb} = X_4$.

\begin{equation}
-\frac{1}{\gamma}\lambda_{fs} = \big( \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\big) \frac{1}{(Y_{fb} - Y_{fm})}
\end{equation}

\begin{equation}
-\frac{1}{\gamma}\lambda_{fs} = \big( \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr} \big) \frac{1}{(X_{fb} - X_{fm})}
\end{equation}

\begin{equation}
  \big( \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\big) (X_{fb} - X_{fm}) =
  \big( \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr} \big) (Y_{fb} - Y_{fm})
\end{equation}

\begin{equation}
  \big( \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\big) (-L_s\cos{\Theta_{fb}}) =
  \big( \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr} \big) (-L_s\sin{\Theta_{fb}})
\end{equation}

\begin{equation}
  \big( \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb} - R_{ybr} - \frac{1}{\gamma}F_{ybr}\big) (\cos{\Theta_{fb}}) =
  \big( \dot{X}_{fm} + L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb} - R_{xbr} - \frac{1}{\gamma}F_{xbr} \big) (\sin{\Theta_{fb}})
\end{equation}

\begin{equation}
  \big( \dot{Y}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\big)(\cos{\Theta_{fb}}) + \big(-\dot{X}_{fm} - L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}\big)(\sin{\Theta_{fb}}) =
  \big(- R_{xbr} - \frac{1}{\gamma}F_{xbr} \big) (\sin{\Theta_{fb}}) + \big(R_{ybr} + \frac{1}{\gamma}F_{ybr}\big) (\cos{\Theta_{fb}})
\end{equation}

\begin{multline}
  \big(L_{s}\cos(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\cos(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\cos(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\cos(\Theta_{fb})\dot{\Theta}_{fb}\big)(\cos{\Theta_{fb}})\\
+ \big(L_{s}\sin(\Theta_{bb})\dot{\Theta}_{bb} + L_{t}\sin(\Theta_{bm})\dot{\Theta}_{bm} - L_{t}\sin(\Theta_{fm})\dot{\Theta}_{fm} - L_{s}\sin(\Theta_{fb})\dot{\Theta}_{fb}\big)(\sin{\Theta_{fb}}) = \\
  \big(- R_{xbr} - \frac{1}{\gamma}F_{xbr} \big) (\sin{\Theta_{fb}}) + \big(R_{ybr} + \frac{1}{\gamma}F_{ybr}\big) (\cos{\Theta_{fb}})
\end{multline}

\begin{multline}
  L_{s}(\cos(\Theta_{fb})\cos(\Theta_{bb}) + \sin(\Theta_{fb})\sin(\Theta_{bb}))\dot{\Theta}_{bb}
  + L_{t}(\cos(\Theta_{fb})\cos(\Theta_{bm}) + \sin(\Theta_{fb})\sin(\Theta_{bm}))\dot{\Theta}_{bm} \\
  - L_{t}(\cos(\Theta_{fb})\cos(\Theta_{fm}) + \sin(\Theta_{fb})\sin(\Theta_{fm}))\dot{\Theta}_{fm}
  - L_{s}(\cos^2(\Theta_{fb}) + \sin^2(\Theta_{fb}))\dot{\Theta}_{fb} = \\
  (R_{ybr} + \frac{1}{\gamma}F_{ybr}) (\cos{\Theta_{fb}}) - (R_{xbr} + \frac{1}{\gamma}F_{xbr} ) (\sin{\Theta_{fb}})
\end{multline}

\begin{align}
  D_1 &= L_{s}\Big(\cos(\Theta_{fb})\cos(\Theta_{bb}) + \sin(\Theta_{fb})\sin(\Theta_{bb})\Big)\\
  D_2 &= L_{t}\Big(\cos(\Theta_{fb})\cos(\Theta_{bm}) + \sin(\Theta_{fb})\sin(\Theta_{bm})\Big)\\
  D_3 &= -L_{t}\Big(\cos(\Theta_{fb})\cos(\Theta_{fm}) + \sin(\Theta_{fb})\sin(\Theta_{fm})\Big)\\
  D_4 &= -L_{s}\\
  X_4 &= (R_{ybr} + \frac{1}{\gamma}F_{ybr})\cos{\Theta_{fb}} - (R_{xbr} + \frac{1}{\gamma}F_{xbr})\sin(\Theta_{fb})
\end{align}

\section{CAS solution to Matrix Equation}

We now have our equations in the following form:

\begin{equation}
  \begin{pmatrix}
    A_1 & A_2 & A_3 & A_4 \\
    B_1 & B_2 & B_3 & B_4 \\
    C_1 & C_2 & C_3 & C_4 \\
    D_1 & D_2 & D_3 & D_4 \\    
  \end{pmatrix}
\times
  \begin{pmatrix}
    \Theta_{bb} \\
    \Theta_{bm} \\
    \Theta_{fm} \\
    \Theta_{fb} \\
  \end{pmatrix}
=
  \begin{pmatrix}
    X_1 \\
    X_2 \\
    X_3 \\
    X_4 \\
  \end{pmatrix}
\end{equation}

We use a CAS (Mathematica) to solve this equation for the angular velocities. We arrive at the following expressions, which we have broken up into numerator and denominator for legibility:

\begin{align}
\dot{\Theta}_{bb} &= \frac{N_{bb}}{D} \\
\dot{\Theta}_{bm} &= \frac{N_{bm}}{D} \\
\dot{\Theta}_{fm} &= \frac{N_{fm}}{D} \\
\dot{\Theta}_{fb} &= \frac{N_{fb}}{D}
\end{align}

\begin{multline}
N_{bb} =
(-B_2 C_4 D_3 X_1 + B_2 C_3 D_4 X_1 + A_4 C_3 D_2 X_2 - A_3 C_4 D_2 X_2 - A_4 C_2 D_3 X_2 + A_2 C_4 D_3 X_2 +\\
 A_3 C_2 D_4 X_2 - A_2 C_3 D_4 X_2 + A_4 B_2 D_3 X_3 - A_3 B_2 D_4 X_3 - A_4 B_2 C_3 X_4 + A_3 B_2 C_4 X_4 +\\
B_4 (-C_3 D_2 X_1 + C_2 D_3 X_1 + A_3 D_2 X_3 - A_2 D_3 X_3 - A_3 C_2 X_4 + A_2 C_3 X_4) + B_3 (C_4 D_2 X_1 - \\
C_2 D_4 X_1 - A_4 D_2 X_3 + A_2 D_4 X_3 + A_4 C_2 X_4 - A_2 C_4 X_4)) \\
\end{multline}

\begin{multline}
N_{bm} =
(B_1 C_4 D_3 X_1 - B_1 C_3 D_4 X_1 - A_4 C_3 D1 X_2 + A_3 C_4 D1 X_2 + A_4 C_1 D_3 X_2 - A_1 C_4 D_3 X_2 - \\
A_3 C_1 D_4 X_2 + A_1 C_3 D_4 X_2 - A_4 B_1 D_3 X_3 + A_3 B_1 D_4 X_3 + A_4 B_1 C_3 X_4 - A_3 B_1 C_4 X_4 +\\
B_4 (C_3 D1 X_1 - C_1 D_3 X_1 - A_3 D1 X_3 + A_1 D_3 X_3 + A_3 C_1 X_4 - A_1 C_3 X_4) + B_3 (-C_4 D1 X_1 + \\
C_1 D_4 X_1 + A_4 D1 X_3 - A_1 D_4 X_3 - A_4 C_1 X_4 + A_1 C_4 X_4)) \\
\end{multline}

\begin{multline}
N_{fm} =
(-B_1 C_4 D_2 X_1 + B_1 C_2 D_4 X_1 + A_4 C_2 D1 X_2 - A_2 C_4 D1 X_2 - A_4 C_1 D_2 X_2 + A_1 C_4 D_2 X_2 + \\
A_2 C_1 D_4 X_2 - A_1 C_2 D_4 X_2 + A_4 B_1 D_2 X_3 - A_2 B_1 D_4 X_3 - A_4 B_1 C_2 X_4 + A_2 B_1 C_4 X_4 +\\
B_4 (-C_2 D1 X_1 + C_1 D_2 X_1 + A_2 D1 X_3 - A_1 D_2 X_3 - A_2 C_1 X_4 + A_1 C_2 X_4) + \\
B_2 (C_4 D1 X_1 - C_1 D_4 X_1 - A_4 D1 X_3 + A_1 D_4 X_3 + A_4 C_1 X_4 - A_1 C_4 X_4)) \\
\end{multline}

\begin{multline}
N_{fb} =
(B_1 C_3 D_2 X_1 - B_1 C_2 D_3 X_1 - A_3 C_2 D1 X_2 + A_2 C_3 D1 X_2 + A_3 C_1 D_2 X_2 - A_1 C_3 D_2 X_2 - \\
A_2 C_1 D_3 X_2 + A_1 C_2 D_3 X_2 - A_3 B_1 D_2 X_3 + A_2 B_1 D_3 X_3 + A_3 B_1 C_2 X_4 - A_2 B_1 C_3 X_4 +\\
B_3 (C_2 D1 X_1 - C_1 D_2 X_1 - A_2 D1 X_3 + A_1 D_2 X_3 + A_2 C_1 X_4 - A_1 C_2 X_4) + \\
B_2 (-C_3 D1 X_1 + C_1 D_3 X_1 + A_3 D1 X_3 - A_1 D_3 X_3 - A_3 C_1 X_4 + A_1 C_3 X_4)) \\
\end{multline}

\begin{multline}
D = (A_2 B_4 C_3 D1 - A_2 B_3 C_4 D1 - A_1 B_4 C_3 D_2 + A_1 B_3 C_4 D_2 - A_2 B_4 C_1 D_3 + A_1 B_4 C_2 D_3 + \\
A_2 B_1 C_4 D_3 - A_1 B_2 C_4 D_3 + A_4 (B_3 C_2 D1 - B_2 C_3 D1 - B_3 C_1 D_2 + B_1 C_3 D_2 + B_2 C_1 D_3 - B_1 C_2 D_3) + \\
A_2 B_3 C_1 D_4 - A_1 B_3 C_2 D_4 - A_2 B_1 C_3 D_4 + A_1 B_2 C_3 D_4 + A_3 (-B_4 C_2 D1 + B_2 C_4 D1 + B_4 C_1 D_2 - B_1 C_4 D_2 - B_2 C_1 D_4 + B_1 C_2 D_4))
\end{multline}

\clearpage
{\huge Bothbound Solution} \\
We now do a similar solution for velocities in the case where both of the model's binding domains are stuck to the microtubule.

\begin{figure}[h!]
  \caption{Key lengths and angles in bothbound dynein model.}
  \centering
    \includegraphics[width=0.7\textwidth]{../figures/bothbound_model.png}
\end{figure}

\section{Cartesian Coordinate Derivation}

By the law of cosines:
\begin{align}
  L_n &= \sqrt{L_{s}^2+L_{t}^2-2L_{t}L_{s}\cos(\Theta_{nm})}\\
  L_f &= \sqrt{L_{s}^2+L_{t}^2-2L_{t}L_{s}\cos(\Theta_{fm})}\\
\end{align}

By the law of cosines, we see that:
\begin{align}
  \Theta_{1} &= \cos^{-1}\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg) \\
  \Theta_{2} &= \pm\cos^{-1}\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg) \\
\end{align}

\subsection{Solving for $X_{nm}$ and $Y_{nm}$}

\begin{align}
  X_{nm} &= L_{s}*\cos(\Theta_{1}+\Theta_{2}) =
     L_{s}*\cos\Bigg(\cos^{-1}\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg)
     \pm \cos^{-1}\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg)\Bigg) \\
  Y_{nm} &= L_{s}*\sin(\Theta_{1}+\Theta_{2}) =
     L_{s}*\sin\Bigg(\cos^{-1}\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg)
     \pm \cos^{-1}\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg)\Bigg) \\
\end{align}

We then use the trigonometric rules $$\cos\Big(cos^{-1}(a)\pm
cos^{-1}(b)\Big) = ab \mp \sqrt{1-a^2}\sqrt{1-b^2}$$
and $$\sin\Big(cos^{-1}(a)\pm cos^{-1}(b)\Big) = \sqrt{1-a^2}b\pm a\sqrt{1-b^2}$$ to simplify our expressions:

\begin{align}
  X_{nm} &= L_{s}\Bigg( \Big(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\Big)\Big(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\Big) \mp \sqrt{1-\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg)^2}\sqrt{1-\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg)^2} \Bigg)\\
  Y_{nm} &= L_{s}\Bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\sqrt{1-\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg)^2} \pm \frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\sqrt{1-\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg)^2} \Bigg)\\
\end{align}

We simplify further...

\begin{align}
  X_{nm} &= \Big(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\Big)\Big(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{n}}\Big) \mp L_s\sqrt{1-\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg)^2}\sqrt{1-\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg)^2}\\
  Y_{nm} &= \frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{n}}\sqrt{1-\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg)^2} \pm L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\sqrt{1-\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg)^2} \Bigg)\\
\end{align}

\begin{align}
  X_{nm} &= \frac{1}{4LL_s}\Big(L^2+L_{n}^2-L_{f}^2\Big)\Big(L_{s}^2+L_{n}^2-L_{t}^2\Big)\big(L_n \big)^{-2} \mp L_s\sqrt{1-\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg)^2}\sqrt{1-\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg)^2}\\
  Y_{nm} &= \frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{n}}\sqrt{1-\bigg(\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\bigg)^2} \pm L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\sqrt{1-\bigg(\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{s}L_{n}}\bigg)^2} \Bigg)\\
\end{align}

\begin{align}
  X_{nm} &= \frac{L_s^2L^2 + L_n^2L^2 - L_t^2L^2 + L_s^2L_{n}^2 + L_n^2L_{n}^2 - L_t^2L_{n}^2 - L_s^2L_{f}^2 - L_n^2L_{f}^2 + L_t^2L_{f}^2}{4L_n^2LL_s}\\
  &\mp L_s\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}*\\
  &\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}\\
  Y_{nm} &= \frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{n}} \sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 - L_n^2L_{f}^2 + L_{f}^4}{4L^2L^2_{n}}}\\ &\pm L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\sqrt{1 - \frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L^2_{s}L^2_{n}}}\\
\end{align}

Now we derive with respect to time. Things which are time-dependent: $L_n, L_f$.

\begin{align}
  \dot{X}_{nm} &= \frac{1}{{2L_n^{-3}LL_s}}\left(-L_s^2L^2 - L_n^2L^2 + L_t^2L^2 - L_s^2L_{n}^2 - L_n^2L_{n}^2 + L_t^2L_{n}^2 + L_s^2L_{f}^2 + L_n^2L_{f}^2 - L_t^2L_{f}^2\right)\dot{L}_n\\ &+ \frac{2L_nL^2\dot{L}_n + 2L_s^2L_{n}\dot{L}_n + 4L_n^3\dot{L}_n - 2L_t^2L_{n}\dot{L}_n - 2L_s^2L_{f}\dot{L}_f - 2L_nL_{f}^2\dot{L}_n - 2L_n^2L_{f}\dot{L}_f + 2L_t^2L_{f}\dot{L}_f}{4L_n^2LL_s}\\
  &\mp L_s\frac{\frac{d}{dt}\left(1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}\right)}{2\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}}*\\
  &\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}\\
  &\mp L_s\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}*\\
  &\frac{\frac{d}{dt}\left(1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}\right)}{2\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}}\\
\end{align}

\begin{align}
  \dot{X}_{nm} &= \frac{-L_s^2L^2 - L_n^2L^2 + L_t^2L^2 - L_s^2L_{n}^2 - L_n^2L_{n}^2 + L_t^2L_{n}^2 + L_s^2L_{f}^2 + L_n^2L_{f}^2 - L_t^2L_{f}^2}{{2L_n^{3}LL_s}}\dot{L}_n\\ &+ \frac{2L_nL^2\dot{L}_n + 2L_s^2L_{n}\dot{L}_n + 4L_n^3\dot{L}_n - 2L_t^2L_{n}\dot{L}_n - 2L_s^2L_{f}\dot{L}_f - 2L_nL_{f}^2\dot{L}_n - 2L_n^2L_{f}\dot{L}_f + 2L_t^2L_{f}\dot{L}_f}{4L_n^2LL_s}\\
  &\mp L_s\frac{\left(\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL^2_{n}} - \frac{2L_nL^2\dot{L}_n - 2L_fL^2\dot{L}_f + 2L^2L_{n}\dot{L}_n + 4L_{n}^3\dot{L}_n - 2L^2L_{f}\dot{L}_f  + 4L_{f}^3\dot{L}_f}{2LL_{n}}\right)}{2\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}}*\\
  &\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}\\
  &\mp L_s\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}*\\
  &\frac{\left(\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L^2_{n}}\dot{L}_n
    - \frac{2L_nL_{s}^2\dot{L}_n + 2L_s^2L_{n}\dot{L}_n + 4L_{n}^3\dot{L}_n - 2L_t^2L_{n}\dot{L}_n - 2L_nL_{t}^2\dot{L}_n}{2L_{s}L_{n}}\right)}{2\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}}\\
\end{align}

\begin{align}
  \dot{X}_{nm} &= \frac{-L_s^2L^2 - L_n^2L^2 + L_t^2L^2 - L_s^2L_{n}^2 - L_n^2L_{n}^2 + L_t^2L_{n}^2 + L_s^2L_{f}^2 + L_n^2L_{f}^2 - L_t^2L_{f}^2}{{2L_n^{3}LL_s}}\dot{L}_n\\
  &+ \frac{2L_nL^2 + 2L_s^2L_{n} + 4L_n^3 - 2L_t^2L_{n} - 2L_nL_{f}^2\dot{L}_n}{4L_n^2LL_s}\dot{L}_n + \frac{-2L_s^2L_{f} - 2L_n^2L_{f} + 2L_t^2L_{f}}{4L_n^2LL_s}\dot{L}_f\\
  &\mp L_s\frac{\left(\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL^2_{n}} - \frac{2L_nL^2 + 2L^2L_{n} + 4L_{n}^3}{2LL_{n}}\right)}{2\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}}*\\
  &\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}\dot{L}_n\\
  &\mp L_s\frac{\left(-\frac{-2L_fL^2 - 2L^2L_{f} + 4L_{f}^3}{2LL_{n}}\right)}{2\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}}*\\
  &\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}\dot{L}_f\\
  &\mp L_s\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}*\\
  &\frac{\left(\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L^2_{n}}
    - \frac{2L_nL_{s}^2 + 2L_s^2L_{n} + 4L_{n}^3 - 2L_t^2L_{n} - 2L_nL_{t}^2}{2L_{s}L_{n}}\right)}{2\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}}\dot{L}_n\\
\end{align}

\begin{align}
  \dot{X}_{nm} &= \Bigg(\frac{-L_s^2L^2 - L_n^2L^2 + L_t^2L^2 - L_s^2L_{n}^2 - L_n^2L_{n}^2 + L_t^2L_{n}^2 + L_s^2L_{f}^2 + L_n^2L_{f}^2 - L_t^2L_{f}^2}{{2L_n^{3}L}}\\
  &+ \frac{2L_nL^2 + 2L_s^2L_{n} + 4L_n^3 - 2L_t^2L_{n} - 2L_nL_{f}^2}{4L_n^2L}\\
  &\mp \frac{\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L_{s}^2L_{n}^2}}}{2\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{4L^2L_{n}^2}}}*\\
  &\left(L_s\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{4L^2L^3_{n}} - L_s\frac{2L_nL^2 + 2L^2L_{n} + 4L_{n}^3}{4L^2L_{n}^2}\right)\\
  &\mp \left(L_s\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L_{s}^2L^3_{n}}
    - L_s\frac{2L_nL_{s}^2 + 2L_s^2L_{n} + 4L_{n}^3 - 2L_t^2L_{n} - 2L_nL_{t}^2}{4L_{s}^2L_{n}^2}\right)\\
  &\frac{\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}}{2\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}}\Bigg)\dot{L}_n\\
  &\Bigg(\frac{-2L_fL_s^2 - 2L_fL_n^2 + 2L_fL_t^2}{4LL_{n}^2}
  \mp L_s\frac{\sqrt{1-\left(\frac{L_s^4 + L_s^2L_n^2 - L_s^2L_t^2 L_n^2L^2 + L_n^4 - L_n^2L_f^2 - L_f^2L^2 - L_f^2L_n^2 + L_f^4}{4L_{s}^2L_{n}^2}\right)}}{2\sqrt{1 - \left(\frac{L^4 + L^2L_n^2 - L^2L_f^2 + L_n^2L^2 + L_n^4 - L_n^2L_f^2 - L_f^2L^2 - L_f^2L_n^2 + L_f^4}{4L^2L_{n}^2}\right)}}
  *\left(\frac{2L^2L_f - 2L_n^2L_f - 2L_fL^2 - 2L_fL_n^2 + 4L_f^3}{4L^2L_{n}^2}\right)\\
  &\mp L_s\frac{\sqrt{1 - \left(\frac{L^4 + L^2L_n^2 - L^2L_f^2 + L_n^2L^2 + L_n^4 - L_n^2L_f^2 - L_f^2L^2 - L_f^2L_n^2 + L_f^4}{4L^2L_{n}^2}\right)}}{2\sqrt{1-\left(\frac{L_s^4 + L_s^2L_n^2 - L_s^2L_t^2 L_n^2L^2 + L_n^4 - L_n^2L_f^2 - L_f^2L^2 - L_f^2L_n^2 + L_f^4}{4L_{s}^2L_{n}^2}\right)}}
  *\left(\frac{-2L_n^2L_f - 2L_fL^2 - 2L_fL_n^2 + 4L_f^3}{4L_{s}^2L_{n}^2}\right)\Bigg)\dot{L}_f\\
    %% &+ \Bigg(\frac{-L_s^2L_{f} - L_n^2L_{f} + L_t^2L_{f}}{2L_n^2L} \mp L_s\frac{\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}}{2\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}}*\\
  %% &\left(-\frac{-2L_fL^2 - 2L^2L_{f} + 4L_{f}^3}{2LL_{n}}\right)\Bigg)\dot{L}_f\\
\end{align}

We now do the same for $\dot{Y}_{nm}$.

\begin{align}
  Y_{nm} &= \frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L^2_{n}} \sqrt{L_n^2-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 - L_n^2L_{f}^2 + L_{f}^4}{4L^2}}\\
  &\pm L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL^2_{n}}\sqrt{L^2_n - \frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L^2_{s}}}\\
\end{align}

\begin{align}
  \dot{Y}_{nm} &= \left(\frac{L_{s}^2+2L_{n}\dot{L}_n - L_{t}^2}{2L^2_{n}} - 2\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L^3_{n}}\dot{L}_n\right) \sqrt{L_n^2-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 - L_n^2L_{f}^2 + L_{f}^4}{4L^2}}\\
  &+ \frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L^2_{n}} \left(\frac{d}{dt} \sqrt{L_n^2-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 - L_n^2L_{f}^2 + L_{f}^4}{4L^2}}\right)\\
  &\pm L_s\left(\frac{L^2+2L_{n}\dot{L}_n - 2L_{f}\dot{L}_f}{2LL^2_n} - 2\frac{L^2+L_{n}^2-L_{f}^2}{2LL^3_n}\dot{L}_n\right) \sqrt{L^2_n - \frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L^2_{s}}}\\  
  &\pm L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL^2_{n}} \left(\frac{d}{dt} \sqrt{L^2_n - \frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L^2_{s}}}\right)\\
\end{align}

\begin{align}
  \dot{Y}_{nm} &= \left(\frac{L_{s}^2+2L_{n}\dot{L}_n - L_{t}^2}{2L^2_{n}} - 2\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L^3_{n}}\dot{L}_n\right) \sqrt{L_n^2-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 - L_n^2L_{f}^2 + L_{f}^4}{4L^2}}\\
  &+ \frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L^2_{n}}   \frac{2L_n\dot{L}_n - \frac{L^4 + 2L_nL^2\dot{L}_n - 2L_fL^2\dot{L}_f + 2L^2L_{n}\dot{L}_n + 4L_{n}^3\dot{L}_n - 2L_fL_{n}^2\dot{L}_f - 2L_f^2L_{n}\dot{L}_n - 2L^2L_{f}\dot{L}_f - 2L_nL_{f}^2\dot{L}_n - 2L_n^2L_{f}\dot{L}_f + 4L_{f}^3\dot{L}_f}{4L^2}}{2\sqrt{L_n^2-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 - L_n^2L_{f}^2 + L_{f}^4}{4L^2}}}\\
  &\pm L_s\left(\frac{L^2+2L_{n}\dot{L}_n - 2L_{f}\dot{L}_f}{2LL^2_n} - 2\frac{L^2+L_{n}^2-L_{f}^2}{2LL^3_n}\dot{L}_n\right) \sqrt{L^2_n - \frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L^2_{s}}}\\  
  &\pm L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL^2_{n}} \left(\frac{2L_n\dot{L}_n - \frac{2L_nL_{s}^2\dot{L}_n + 2L_s^2L_{n}\dot{L}_n + 4L_{n}^3\dot{L}_n - 2L_t^2L_{n}\dot{L}_n - 2L_nL_{t}^2\dot{L}_n}{4L^2_{s}}}{2\sqrt{L^2_n - \frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L^2_{s}}}} \right)\\
\end{align}

\begin{align}
  \dot{Y}_{nm} &= \left(\frac{L_{s}^2+2L_{n}\dot{L}_n - L_{t}^2}{2L^2_{n}} - 2\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L^3_{n}}\dot{L}_n\right) \sqrt{L_n^2-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 - L_n^2L_{f}^2 + L_{f}^4}{4L^2}}\\
  &+ \frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L^2_{n}}   \frac{2L_n\dot{L}_n - \frac{L^4 + 2L_nL^2\dot{L}_n - 2L_fL^2\dot{L}_f + 2L^2L_{n}\dot{L}_n + 4L_{n}^3\dot{L}_n - 2L_fL_{n}^2\dot{L}_f - 2L_f^2L_{n}\dot{L}_n - 2L^2L_{f}\dot{L}_f - 2L_nL_{f}^2\dot{L}_n - 2L_n^2L_{f}\dot{L}_f + 4L_{f}^3\dot{L}_f}{4L^2}}{2\sqrt{L_n^2-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 - L_n^2L_{f}^2 + L_{f}^4}{4L^2}}}\\
  &\pm L_s\left(\frac{L^2+2L_{n}\dot{L}_n - 2L_{f}\dot{L}_f}{2LL^2_n} - 2\frac{L^2+L_{n}^2-L_{f}^2}{2LL^3_n}\dot{L}_n\right) \sqrt{L^2_n - \frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L^2_{s}}}\\  
  &\pm L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL^2_{n}} \left(\frac{2L_n\dot{L}_n - \frac{2L_nL_{s}^2\dot{L}_n + 2L_s^2L_{n}\dot{L}_n + 4L_{n}^3\dot{L}_n - 2L_t^2L_{n}\dot{L}_n - 2L_nL_{t}^2\dot{L}_n}{4L^2_{s}}}{2\sqrt{L^2_n - \frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L^2_{s}}}} \right)\\
\end{align}

\begin{align}
    \dot{Y}_{nm} &=
    \Bigg(\left(-\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_n^2} + 1\right)\sqrt{1-\frac{L^4 + L_n^4 + L_f^4 + 2L^2L_n^2 - 2L^2L_f^2 - 2L_n^2L_f^2}{4L^2L_n^2}} +\\
    &\frac{-2\frac{L^4 + L_n^4 + L_f^4 + 2L^2L_n^2 - 2L^2L_f^2 - 2L_n^2L_f^2}{4L^2L_n^3} + \frac{4L_n^3 + 4L^2L_n - 4L_n^2L_f}{4L^2L_n^2}}{2\sqrt{1-\frac{L^4 + L_n^4 + L_f^4 + 2L^2L_n^2 - 2L^2L_f^2 - 2L_n^2L_f^2}{4L^2L_n^2}}}\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{n}} \pm\\
    &\left(-L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}^2} + \frac{L_s\dot{L}_n}{L}\right)\sqrt{1-\frac{L_s^4 + L_n^4 + L_t^4 + 2L_sL_n^2 - 2L_s^2L_t^2 - 2L_n^2L_t^2}{4L_s^2L_n^2}} \pm\\
    &L_s\frac{-2\frac{L_s^4 + L_n^4 + L_t^4 + 2L_sL_n^2 - 2L_s^2L_t^2 - 2L_n^2L_t^2}{4L_s^2L_n^3} + \frac{4L_n^3 + 4L_sL_n - 4L_nL_t^2}{4L_s^2L_n^2}}{2\sqrt{1-\frac{L_s^4 + L_n^4 + L_t^4 + 2L_sL_n^2 - 2L_s^2L_t^2 - 2L_n^2L_t^2}{4L_s^2L_n^2}}}\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\Bigg)\dot{L}_n\\
    + &\Bigg(\frac{\frac{4L_f^3 - 4L^2L_f - 4L_n^2L_f}{4L^2L_n^2}}{2\sqrt{1-\frac{L^4 + L_n^4 + L_f^4 + 2L^2L_n^2 - 2L^2L_f^2 - 2L_n^2L_f^2}{4L^2L_n^2}}}\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{n}}\\
    \pm &\left(\frac{-L_sL_{f}}{LL_{n}}\right)\sqrt{1-\frac{L_s^4 + L_n^4 + L_t^4 + 2L_sL_n^2 - 2L_s^2L_t^2 - 2L_n^2L_t^2}{4L_s^2L_n^2}}\Bigg)\dot{L}_f\\
\end{align}

\subsection{Solving for $X_{fm}$ and $Y_{fm}$}
The nearbound and farbound angles are symmetric. We switch all instances of $L_n$ and $L_f$.

\begin{align}
  X_{fm} &= \frac{1}{4LL_s}\Big(L^2+L_{f}^2-L_{n}^2\Big)\Big(L_{s}^2+L_{f}^2-L_{t}^2\Big)\big(L_f \big)^{-2} \mp L_s\sqrt{1-\bigg(\frac{L^2+L_{f}^2-L_{n}^2}{2LL_{f}}\bigg)^2}\sqrt{1-\bigg(\frac{L_{s}^2+L_{f}^2-L_{t}^2}{2L_{s}L_{f}}\bigg)^2}\\
  Y_{fm} &= \frac{L_{s}^2+L_{f}^2-L_{t}^2}{2L_{f}}\sqrt{1-\bigg(\frac{L^2+L_{f}^2-L_{n}^2}{2LL_{f}}\bigg)^2} \pm L_s\frac{L^2+L_{f}^2-L_{n}^2}{2LL_{f}}\sqrt{1-\bigg(\frac{L_{s}^2+L_{f}^2-L_{t}^2}{2L_{s}L_{f}}\bigg)^2} \Bigg)\\
\end{align}

\begin{align}
  \dot{X}_{fm} &= \Bigg(\frac{-L_s^2L^2 - L_f^2L^2 + L_t^2L^2 - L_s^2L_f^2 - L_f^2L_f^2 + L_t^2L_f^2 + L_s^2L_n^2 + L_f^2L_n^2 - L_t^2L_n^2}{{2L_f^{3}L}}\\
  &+ \frac{2L_fL^2 + 2L_s^2L_f + 4L_f^3 - 2L_t^2L_f - 2L_fL_n^2}{4L_f^2L}\\
  &\mp \frac{\sqrt{1-\frac{L_{s}^4 + L_f^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_f^2 + L_f^4 - L_t^2L_f^2 - L_s^2L_{t}^2 - L_f^2L_{t}^2 + L_{t}^4}{4L_{s}^2L_f^2}}}{2\sqrt{1-\frac{L^4 + L_f^2L^2 - L_n^2L^2 + L^2L_f^2 + L_f^4 - L_n^2L_f^2 - L^2L_n^2 + L_f^2L_n^2 + L_n^4}{4L^2L_f^2}}}*\\
  &\left(L_s\frac{L^4 + L_f^2L^2 - L_n^2L^2 + L^2L_f^2 + L_f^4 - L_n^2L_f^2 - L^2L_n^2 + L_f^2L_n^2 + L_n^4}{4L^2L^3_{n}} - L_s\frac{2L_fL^2 + 2L^2L_f + 4L_f^3}{4L^2L_f^2}\right)\\
  &\mp \left(L_s\frac{L_{s}^4 + L_f^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_f^2 + L_f^4 - L_t^2L_f^2 - L_s^2L_{t}^2 - L_f^2L_{t}^2 + L_{t}^4}{4L_{s}^2L^3_{n}}
    - L_s\frac{2L_fL_{s}^2 + 2L_s^2L_f + 4L_f^3 - 2L_t^2L_f - 2L_fL_{t}^2}{4L_{s}^2L_f^2}\right)\\
  &\frac{\sqrt{1-\frac{L^4 + L_f^2L^2 - L_n^2L^2 + L^2L_f^2 + L_f^4 - L_n^2L_f^2 - L^2L_n^2 + L_f^2L_n^2 + L_n^4}{2LL_f}}}{2\sqrt{1-\frac{L_{s}^4 + L_f^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_f^2 + L_f^4 - L_t^2L_f^2 - L_s^2L_{t}^2 - L_f^2L_{t}^2 + L_{t}^4}{2L_{s}L_f}}}\Bigg)\dot{L}_f\\
  &\Bigg(\frac{-2L_nL_s^2 - 2L_nL_f^2 + 2L_nL_t^2}{4LL_f^2}
  \mp L_s\frac{\sqrt{1-\left(\frac{L_s^4 + L_s^2L_f^2 - L_s^2L_t^2 L_f^2L^2 + L_f^4 - L_f^2L_n^2 - L_n^2L^2 - L_n^2L_f^2 + L_n^4}{4L_{s}^2L_f^2}\right)}}{2\sqrt{1 - \left(\frac{L^4 + L^2L_f^2 - L^2L_n^2 + L_f^2L^2 + L_f^4 - L_f^2L_n^2 - L_n^2L^2 - L_n^2L_f^2 + L_n^4}{4L^2L_f^2}\right)}}
  *\left(\frac{2L^2L_n - 2L_f^2L_n - 2L_nL^2 - 2L_nL_f^2 + 4L_n^3}{4L^2L_f^2}\right)\\
  &\mp L_s\frac{\sqrt{1 - \left(\frac{L^4 + L^2L_f^2 - L^2L_n^2 + L_f^2L^2 + L_f^4 - L_f^2L_n^2 - L_n^2L^2 - L_n^2L_f^2 + L_n^4}{4L^2L_f^2}\right)}}{2\sqrt{1-\left(\frac{L_s^4 + L_s^2L_f^2 - L_s^2L_t^2 L_f^2L^2 + L_f^4 - L_f^2L_n^2 - L_n^2L^2 - L_n^2L_f^2 + L_n^4}{4L_{s}^2L_f^2}\right)}}
  *\left(\frac{-2L_f^2L_n - 2L_nL^2 - 2L_nL_f^2 + 4L_n^3}{4L_{s}^2L_f^2}\right)\Bigg)\dot{L}_n\\
    %% &+ \Bigg(\frac{-L_s^2L_n - L_f^2L_n + L_t^2L_n}{2L_f^2L} \mp L_s\frac{\sqrt{1-\frac{L_{s}^4 + L_f^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_f^2 + L_f^4 - L_t^2L_f^2 - L_s^2L_{t}^2 - L_f^2L_{t}^2 + L_{t}^4}{2L_{s}L_f}}}{2\sqrt{1-\frac{L^4 + L_f^2L^2 - L_n^2L^2 + L^2L_f^2 + L_f^4 - L_n^2L_f^2 - L^2L_n^2 + L_f^2L_n^2 + L_n^4}{2LL_f}}}*\\
  %% &\left(-\frac{-2L_nL^2 - 2L^2L_n + 4L_n^3}{2LL_f}\right)\Bigg)\dot{L}_f\\
\end{align}

\begin{align}
    \dot{Y}_{fm} &=
    \Bigg(\left(-\frac{L_{s}^2+L_f^2-L_{t}^2}{2L_f^2} + 1\right)\sqrt{1-\frac{L^4 + L_f^4 + L_n^4 + 2L^2L_f^2 - 2L^2L_n^2 - 2L_f^2L_n^2}{4L^2L_f^2}} +\\
    &\frac{-2\frac{L^4 + L_f^4 + L_n^4 + 2L^2L_f^2 - 2L^2L_n^2 - 2L_f^2L_n^2}{4L^2L_f^3} + \frac{4L_f^3 + 4L^2L_f - 4L_f^2L_n}{4L^2L_f^2}}{2\sqrt{1-\frac{L^4 + L_f^4 + L_n^4 + 2L^2L_f^2 - 2L^2L_n^2 - 2L_f^2L_n^2}{4L^2L_f^2}}}\frac{L_{s}^2+L_f^2-L_{t}^2}{2L_f} \pm\\
    &\left(-L_s\frac{L^2+L_f^2-L_n^2}{2LL_f^2} + \frac{L_s\dot{L}_n}{L}\right)\sqrt{1-\frac{L_s^4 + L_f^4 + L_t^4 + 2L_sL_f^2 - 2L_s^2L_t^2 - 2L_f^2L_t^2}{4L_s^2L_f^2}} \pm\\
    &L_s\frac{-2\frac{L_s^4 + L_f^4 + L_t^4 + 2L_sL_f^2 - 2L_s^2L_t^2 - 2L_f^2L_t^2}{4L_s^2L_f^3} + \frac{4L_f^3 + 4L_sL_f - 4L_fL_t^2}{4L_s^2L_f^2}}{2\sqrt{1-\frac{L_s^4 + L_f^4 + L_t^4 + 2L_sL_f^2 - 2L_s^2L_t^2 - 2L_f^2L_t^2}{4L_s^2L_f^2}}}\frac{L^2+L_f^2-L_n^2}{2LL_f}\Bigg)\dot{L}_f\\
    + &\Bigg(\frac{\frac{4L_n^3 - 4L^2L_n - 4L_f^2L_n}{4L^2L_f^2}}{2\sqrt{1-\frac{L^4 + L_f^4 + L_n^4 + 2L^2L_f^2 - 2L^2L_n^2 - 2L_f^2L_n^2}{4L^2L_f^2}}}\frac{L_{s}^2+L_f^2-L_{t}^2}{2L_f}\\
    \pm &\left(\frac{-L_sL_n}{LL_f}\right)\sqrt{1-\frac{L_s^4 + L_f^4 + L_t^4 + 2L_sL_f^2 - 2L_s^2L_t^2 - 2L_f^2L_t^2}{4L_s^2L_f^2}}\Bigg)\dot{L}_n\\
\end{align}

\subsection{Solving for $X_t$ and $Y_t$}

\begin{align}
  X_{t} &= L_n\cos(\Theta_{1}) \\
  &= \frac{L^2+L_n^2-L_f^2}{2L} \\
  \dot{X}_t &= \frac{L_n}{L}\dot{L}_n - \frac{L_f}{L}\dot{L}_f\\
\end{align}

\begin{align}
  Y_{t} &= L_n\sin(\Theta_{1}) \\
  &= L_n\sin\left(\cos^{-1}\left(\frac{L^2+L_n^2-L_f^2}{2LL_n}\right)\right) \\
  &= L_n\sqrt{1 - \left(\frac{L^2+L_n^2-L_f^2}{2LL_n}\right)^2} \\
  &= \sqrt{L_n^2 - \left(\frac{L^2+L_n^2-L_f^2}{2L}\right)^2} \\
  \dot{Y}_t &= \frac{2L_n\dot{L}_n - \left(\frac{L^2+L_n^2-L_f^2}{L}\right)\left(\frac{L_n\dot{L}_n - L_f\dot{L}_f}{L}\right)}{2\sqrt{L_n^2 - \left(\frac{L^2+L_n^2-L_f^2}{2L}\right)^2}}\\
  \dot{Y}_t &= \frac{2L_n - \left(\frac{L^2+L_n^2-L_f^2}{L}\right)\left(\frac{L_n}{L}\right)}{2\sqrt{L_n^2 - \left(\frac{L^2+L_n^2-L_f^2}{2L}\right)^2}}\dot{L}_n +
  \frac{\left(\frac{L^2+L_n^2-L_f^2}{L}\right)\left(\frac{L_f}{L}\right)}{2\sqrt{L_n^2 - \left(\frac{L^2+L_n^2-L_f^2}{2L}\right)^2}}\dot{L}_f\\
\end{align}

\section{Cartesian from Angular Coordinates}
This section lays out the definitions of the coordinates of the Dynein model. \\
\begin{align}
  X_{nb} &= 0 \\
  X_{nm} &=  \\
  X_{t}  &= \frac{L^2+L_n^2-L_f^2}{2L} \\
  X_{fm} &=  \\
  X_{fb} &= L \\
\end{align}

\begin{align}
  Y_{nb} &= 0 \\
  Y_{nm} &=  \\
  Y_{t}  &= L_n\sqrt{1 - \left(\frac{(L^2+L_n^2-L_f^2}{2LL_n}\right)^2} \\
  Y_{fm} &=  \\
  Y_{fb} &= 0
\end{align}

\section{Cartesian Velocities}
This section shows the velocity definitions. \\

\subsection{Velocity Coefficients}
This subsection defines the coeffients to be used in further derivation.\\

\begin{align}
  \dot{X}_{nb} &= 0 \\
  \dot{X}_{nm} &= A_{xn}\dot{L}_n + A_{xf}\dot{L}_f\\
  \dot{X}_{t } &= B_{xn}\dot{L}_n + B_{xf}\dot{L}_f\\
  \dot{X}_{fm} &= C_{xn}\dot{L}_n + C_{xf}\dot{L}_f\\
  \dot{X}_{fb} &= 0
\end{align}

\begin{align}                                                                          
  \dot{Y}_{nb} &= 0 \\                           
  \dot{Y}_{nm} &= A_{yn}\dot{L}_n + A_{yf}\dot{L}_f\\
  \dot{Y}_{t}  &= B_{yn}\dot{L}_n + B_{yf}\dot{L}_f\\
  \dot{Y}_{fm} &= C_{yn}\dot{L}_n + C_{yf}\dot{L}_f\\
  \dot{Y}_{fb} &= 0 \\
\end{align}

\begin{align}
  A_{xn} &= \Bigg(\frac{-L_s^2L^2 - L_n^2L^2 + L_t^2L^2 - L_s^2L_{n}^2 - L_n^2L_{n}^2 + L_t^2L_{n}^2 + L_s^2L_{f}^2 + L_n^2L_{f}^2 - L_t^2L_{f}^2}{{2L_n^{3}L}}\\
  &+ \frac{2L_nL^2 + 2L_s^2L_{n} + 4L_n^3 - 2L_t^2L_{n} - 2L_nL_{f}^2}{4L_n^2L}\\
  &\mp \frac{\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L_{s}^2L_{n}^2}}}{2\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{4L^2L_{n}^2}}}*\\
  &\left(L_s\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{4L^2L^3_{n}} - L_s\frac{2L_nL^2 + 2L^2L_{n} + 4L_{n}^3}{4L^2L_{n}^2}\right)\\
  &\mp \left(L_s\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{4L_{s}^2L^3_{n}}
    - L_s\frac{2L_nL_{s}^2 + 2L_s^2L_{n} + 4L_{n}^3 - 2L_t^2L_{n} - 2L_nL_{t}^2}{4L_{s}^2L_{n}^2}\right)\\
    &\frac{\sqrt{1-\frac{L^4 + L_n^2L^2 - L_f^2L^2 + L^2L_{n}^2 + L_{n}^4 - L_f^2L_{n}^2 - L^2L_{f}^2 + L_n^2L_{f}^2 + L_{f}^4}{2LL_{n}}}}{2\sqrt{1-\frac{L_{s}^4 + L_n^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_{n}^2 + L_{n}^4 - L_t^2L_{n}^2 - L_s^2L_{t}^2 - L_n^2L_{t}^2 + L_{t}^4}{2L_{s}L_{n}}}}\Bigg)\\
  A_{xf} &= \Bigg(\frac{-2L_fL_s^2 - 2L_fL_n^2 + 2L_fL_t^2}{4LL_{n}^2}
  \mp L_s\frac{\sqrt{1-\left(\frac{L_s^4 + L_s^2L_n^2 - L_s^2L_t^2 L_n^2L^2 + L_n^4 - L_n^2L_f^2 - L_f^2L^2 - L_f^2L_n^2 + L_f^4}{4L_{s}^2L_{n}^2}\right)}}{2\sqrt{1 - \left(\frac{L^4 + L^2L_n^2 - L^2L_f^2 + L_n^2L^2 + L_n^4 - L_n^2L_f^2 - L_f^2L^2 - L_f^2L_n^2 + L_f^4}{4L^2L_{n}^2}\right)}}
  *\left(\frac{2L^2L_f - 2L_n^2L_f - 2L_fL^2 - 2L_fL_n^2 + 4L_f^3}{4L^2L_{n}^2}\right)\\
  &\mp L_s\frac{\sqrt{1 - \left(\frac{L^4 + L^2L_n^2 - L^2L_f^2 + L_n^2L^2 + L_n^4 - L_n^2L_f^2 - L_f^2L^2 - L_f^2L_n^2 + L_f^4}{4L^2L_{n}^2}\right)}}{2\sqrt{1-\left(\frac{L_s^4 + L_s^2L_n^2 - L_s^2L_t^2 L_n^2L^2 + L_n^4 - L_n^2L_f^2 - L_f^2L^2 - L_f^2L_n^2 + L_f^4}{4L_{s}^2L_{n}^2}\right)}}
  *\left(\frac{-2L_n^2L_f - 2L_fL^2 - 2L_fL_n^2 + 4L_f^3}{4L_{s}^2L_{n}^2}\right)\Bigg)\\
  B_{xn} &= \frac{L_n}{L}\\
  B_{xf} &= - \frac{L_f}{L}\\
\end{align}
\begin{align}
  C_{xn} &= \Bigg(\frac{-2L_nL_s^2 - 2L_nL_f^2 + 2L_nL_t^2}{4LL_f^2}
  \mp L_s\frac{\sqrt{1-\left(\frac{L_s^4 + L_s^2L_f^2 - L_s^2L_t^2 L_f^2L^2 + L_f^4 - L_f^2L_n^2 - L_n^2L^2 - L_n^2L_f^2 + L_n^4}{4L_{s}^2L_f^2}\right)}}{2\sqrt{1 - \left(\frac{L^4 + L^2L_f^2 - L^2L_n^2 + L_f^2L^2 + L_f^4 - L_f^2L_n^2 - L_n^2L^2 - L_n^2L_f^2 + L_n^4}{4L^2L_f^2}\right)}}
  *\left(\frac{2L^2L_n - 2L_f^2L_n - 2L_nL^2 - 2L_nL_f^2 + 4L_n^3}{4L^2L_f^2}\right)\\
  &\mp L_s\frac{\sqrt{1 - \left(\frac{L^4 + L^2L_f^2 - L^2L_n^2 + L_f^2L^2 + L_f^4 - L_f^2L_n^2 - L_n^2L^2 - L_n^2L_f^2 + L_n^4}{4L^2L_f^2}\right)}}{2\sqrt{1-\left(\frac{L_s^4 + L_s^2L_f^2 - L_s^2L_t^2 L_f^2L^2 + L_f^4 - L_f^2L_n^2 - L_n^2L^2 - L_n^2L_f^2 + L_n^4}{4L_{s}^2L_f^2}\right)}}
  *\left(\frac{-2L_f^2L_n - 2L_nL^2 - 2L_nL_f^2 + 4L_n^3}{4L_{s}^2L_f^2}\right)\Bigg)\\
  C_{xf} &= \Bigg(\frac{-L_s^2L^2 - L_f^2L^2 + L_t^2L^2 - L_s^2L_f^2 - L_f^2L_f^2 + L_t^2L_f^2 + L_s^2L_n^2 + L_f^2L_n^2 - L_t^2L_n^2}{{2L_f^{3}L}}\\
  &+ \frac{2L_fL^2 + 2L_s^2L_f + 4L_f^3 - 2L_t^2L_f - 2L_fL_n^2}{4L_f^2L}\\
  &\mp \frac{\sqrt{1-\frac{L_{s}^4 + L_f^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_f^2 + L_f^4 - L_t^2L_f^2 - L_s^2L_{t}^2 - L_f^2L_{t}^2 + L_{t}^4}{4L_{s}^2L_f^2}}}{2\sqrt{1-\frac{L^4 + L_f^2L^2 - L_n^2L^2 + L^2L_f^2 + L_f^4 - L_n^2L_f^2 - L^2L_n^2 + L_f^2L_n^2 + L_n^4}{4L^2L_f^2}}}*\\
  &\left(L_s\frac{L^4 + L_f^2L^2 - L_n^2L^2 + L^2L_f^2 + L_f^4 - L_n^2L_f^2 - L^2L_n^2 + L_f^2L_n^2 + L_n^4}{4L^2L^3_{n}} - L_s\frac{2L_fL^2 + 2L^2L_f + 4L_f^3}{4L^2L_f^2}\right)\\
  &\mp \left(L_s\frac{L_{s}^4 + L_f^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_f^2 + L_f^4 - L_t^2L_f^2 - L_s^2L_{t}^2 - L_f^2L_{t}^2 + L_{t}^4}{4L_{s}^2L^3_{n}}
    - L_s\frac{2L_fL_{s}^2 + 2L_s^2L_f + 4L_f^3 - 2L_t^2L_f - 2L_fL_{t}^2}{4L_{s}^2L_f^2}\right)\\
    &\frac{\sqrt{1-\frac{L^4 + L_f^2L^2 - L_n^2L^2 + L^2L_f^2 + L_f^4 - L_n^2L_f^2 - L^2L_n^2 + L_f^2L_n^2 + L_n^4}{2LL_f}}}{2\sqrt{1-\frac{L_{s}^4 + L_f^2L_{s}^2 - L_t^2L_{s}^2 + L_s^2L_f^2 + L_f^4 - L_t^2L_f^2 - L_s^2L_{t}^2 - L_f^2L_{t}^2 + L_{t}^4}{2L_{s}L_f}}}\Bigg)\\
\end{align}
\begin{align}
  A_{yn} &= \Bigg(\left(-\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_n^2} + 1\right)\sqrt{1-\frac{L^4 + L_n^4 + L_f^4 + 2L^2L_n^2 - 2L^2L_f^2 - 2L_n^2L_f^2}{4L^2L_n^2}} +\\
    &\frac{-2\frac{L^4 + L_n^4 + L_f^4 + 2L^2L_n^2 - 2L^2L_f^2 - 2L_n^2L_f^2}{4L^2L_n^3} + \frac{4L_n^3 + 4L^2L_n - 4L_n^2L_f}{4L^2L_n^2}}{2\sqrt{1-\frac{L^4 + L_n^4 + L_f^4 + 2L^2L_n^2 - 2L^2L_f^2 - 2L_n^2L_f^2}{4L^2L_n^2}}}\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{n}} \pm\\
    &\left(-L_s\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}^2} + \frac{L_s\dot{L}_n}{L}\right)\sqrt{1-\frac{L_s^4 + L_n^4 + L_t^4 + 2L_sL_n^2 - 2L_s^2L_t^2 - 2L_n^2L_t^2}{4L_s^2L_n^2}} \pm\\
    &L_s\frac{-2\frac{L_s^4 + L_n^4 + L_t^4 + 2L_sL_n^2 - 2L_s^2L_t^2 - 2L_n^2L_t^2}{4L_s^2L_n^3} + \frac{4L_n^3 + 4L_sL_n - 4L_nL_t^2}{4L_s^2L_n^2}}{2\sqrt{1-\frac{L_s^4 + L_n^4 + L_t^4 + 2L_sL_n^2 - 2L_s^2L_t^2 - 2L_n^2L_t^2}{4L_s^2L_n^2}}}\frac{L^2+L_{n}^2-L_{f}^2}{2LL_{n}}\Bigg)\\
  A_{yf} &= \Bigg(\frac{\frac{4L_f^3 - 4L^2L_f - 4L_n^2L_f}{4L^2L_n^2}}{2\sqrt{1-\frac{L^4 + L_n^4 + L_f^4 + 2L^2L_n^2 - 2L^2L_f^2 - 2L_n^2L_f^2}{4L^2L_n^2}}}\frac{L_{s}^2+L_{n}^2-L_{t}^2}{2L_{n}}\\
    \pm &\left(\frac{-L_sL_{f}}{LL_{n}}\right)\sqrt{1-\frac{L_s^4 + L_n^4 + L_t^4 + 2L_sL_n^2 - 2L_s^2L_t^2 - 2L_n^2L_t^2}{4L_s^2L_n^2}}\Bigg)\\
  B_{yn} &= \frac{2L_n - \left(\frac{L^2+L_n^2-L_f^2}{L}\right)\left(\frac{L_n}{L}\right)}{2\sqrt{L_n^2 - \left(\frac{L^2+L_n^2-L_f^2}{2L}\right)^2}}\\
  B_{yf} &= \frac{\left(\frac{L^2+L_n^2-L_f^2}{L}\right)\left(\frac{L_f}{L}\right)}{2\sqrt{L_n^2 - \left(\frac{L^2+L_n^2-L_f^2}{2L}\right)^2}}\\
\end{align}
\begin{align}
  C_{yn} &= \Bigg(\frac{\frac{4L_n^3 - 4L^2L_n - 4L_f^2L_n}{4L^2L_f^2}}{2\sqrt{1-\frac{L^4 + L_f^4 + L_n^4 + 2L^2L_f^2 - 2L^2L_n^2 - 2L_f^2L_n^2}{4L^2L_f^2}}}\frac{L_{s}^2+L_f^2-L_{t}^2}{2L_f}\\
    \pm &\left(\frac{-L_sL_n}{LL_f}\right)\sqrt{1-\frac{L_s^4 + L_f^4 + L_t^4 + 2L_sL_f^2 - 2L_s^2L_t^2 - 2L_f^2L_t^2}{4L_s^2L_f^2}}\Bigg)\\
  C_{yf} &= \Bigg(\left(-\frac{L_{s}^2+L_f^2-L_{t}^2}{2L_f^2} + 1\right)\sqrt{1-\frac{L^4 + L_f^4 + L_n^4 + 2L^2L_f^2 - 2L^2L_n^2 - 2L_f^2L_n^2}{4L^2L_f^2}} +\\
    &\frac{-2\frac{L^4 + L_f^4 + L_n^4 + 2L^2L_f^2 - 2L^2L_n^2 - 2L_f^2L_n^2}{4L^2L_f^3} + \frac{4L_f^3 + 4L^2L_f - 4L_f^2L_n}{4L^2L_f^2}}{2\sqrt{1-\frac{L^4 + L_f^4 + L_n^4 + 2L^2L_f^2 - 2L^2L_n^2 - 2L_f^2L_n^2}{4L^2L_f^2}}}\frac{L_{s}^2+L_f^2-L_{t}^2}{2L_f} \pm\\
    &\left(-L_s\frac{L^2+L_f^2-L_n^2}{2LL_f^2} + \frac{L_s\dot{L}_n}{L}\right)\sqrt{1-\frac{L_s^4 + L_f^4 + L_t^4 + 2L_sL_f^2 - 2L_s^2L_t^2 - 2L_f^2L_t^2}{4L_s^2L_f^2}} \pm\\
    &L_s\frac{-2\frac{L_s^4 + L_f^4 + L_t^4 + 2L_sL_f^2 - 2L_s^2L_t^2 - 2L_f^2L_t^2}{4L_s^2L_f^3} + \frac{4L_f^3 + 4L_sL_f - 4L_fL_t^2}{4L_s^2L_f^2}}{2\sqrt{1-\frac{L_s^4 + L_f^4 + L_t^4 + 2L_sL_f^2 - 2L_s^2L_t^2 - 2L_f^2L_t^2}{4L_s^2L_f^2}}}\frac{L^2+L_f^2-L_n^2}{2LL_f}\Bigg)\\
\end{align}

\begin{twocolumn}
\section{Notes on physical parameters}

Stoke's Law states that the relationship between viscosity, force, and
velocity for a spherical particle is given by:
\begin{align}
  \mathbf{v} &= \frac{\mathbf{F}}{6\pi \mu R}
\end{align}
where $\mu$ is the viscosity of the fluid, and $R$ is the radius of
the particle.  Thus in our terminology:
\begin{align}
  \gamma &= 6\pi \mu R.
\end{align}
Note that the viscosity of water is temperature-dependent and around
0.7~mPa~s at body temperature.  However, the viscosity of cytoplasm
may be rather different, and will of course depend on length scale.
But for our purposes we can probably get away with this value.

The random force $\mathbf{R}(t)$ should have a Gaussian distribution
with variance~1, but then be divided by $\sqrt{\Delta t}$.  Thus, the
smaller the time step, the larger the random force we will apply.  We
can understand this in the backwards direction.  When we observe a
larger time step, the ``random force'' results from many smaller
random forces added together, and these smaller forces often cancel,
leading to this reduction in the total force over that time period.

Finally, we will need estimates for the radii of each domain (to go
into the $\gamma$ parameter, the equilibrium angles for each hinge, and
the spring constants for each hinge.  The first two we can estimate
from the structure of dynein.  The third will require a rough guess.
We can get a ball-park guess by deciding how much we want the angle to
fluctuate (e.g. $\ll \pi$ for each angle!), and then take our spring
constant to be $kT$ divided by that anglular fluctuation squared.

\end{twocolumn}

\end{document}