### 2019-01-22 papers/waczak-thesis/thesis/chapters/dynein.tex
updates to thesis

\section{Directed Motion}
\section{Drunken Walking}
\section{Transporting Cargo}
\section{Experimental Stepping Statistics}