### 2019-01-22 papers/waczak-thesis/thesis/chapters/methods.tex

\section{Theory}
Now that I have established some facts about motor proteins as well as specifics about dynein's many odd behaviors, we may begin our discussion of how to construct and test a physical model for its motion. First, we note that dynein's large mass of roughly 2MDa \cite{johnson_structure_1983} as compared to similar motor proteins such as Kinesin (a few hundred kDa) \cite{liao_kinesin_1998} exaggerates the effect of interactions with the  aqueous environment of the cell. This results in a much more significant drag force as well as a random element from water molecule collisions that changes the overall dynamics of the system. The set of equations governing this motion is called Brownian dynamics and is what we use to simulate the trajectory of dynein along the microtubule. 
	\subsection{Brownian dynamics}
	To model dynein we treat its various domains (binding, motor, and tail) as spheres held together by massless rigid-rods. In general, the dynamics of a constant mass particle are governed by Newton's second law
	\begin{equation}
		\sum_i \mathbf{F}_i = \dot{\mathbf{p}} = m\ddot{\mathbf{x}}
	\end{equation}
	Thus, determining the dynamics of a system boils down to describing the various forces which act upon it and then solving the resulting differential equation. For particles in a solvent, there are three important forces to consider. The first is linear drag, which we write as 
	\begin{equation}
		F_\text{drag} = -\gamma\dot{\mathbf{x}}
	\end{equation}
	where $\gamma$ is the constant of proportionality. For a spherical particle of radius $a$ moving through a solvent with liquid viscosity $\eta$ this was derived by Stoke to be 
	\begin{equation}
		\gamma = 6\pi\eta a
	\end{equation}

	The second characteristic force of Brownian dynamics is a random force due to density fluctuations in the solvent which we denote $F_\text{rand}$. Finally, might also consider how individual particles interact with each other. For us, this will be accomplished through the rigid rods that outline the structure of the protein. Putting all of these forces together results in what is often called the \textit{Langevin Equation}.
	\begin{equation}
		m\ddot{\mathbf{x}} + \gamma\dot{\mathbf{x}}+\mathbf{F}_\text{rand} = 0 
	\end{equation}

	If we assume further that $|\gamma\dot{\mathbf{x}}|\gg |m\ddot{\mathbf{x}}|$, or in other words, we take the limit of strong drag, then we arrive at the definition of Brownian Dynamics. This allows us to neglect the acceleration term resulting in the following differential equation:
	\begin{equation}
		\dot{\mathbf{x}} = \frac{1}{\gamma}\left( \mathbf{F}_\text{interact} + \mathbf{F}_\text{rand} \right)
	\end{equation}
	To complete the model, we must specify how we calculate and apply the random force as well as the force due to the interactions between domains.

	\subsection{A Simplified Model for Dynein}
	As previously mentioned, we model dynein as a system of spherical domains held together by massless rigid-rods with the whole system moving according to Brownian Dynamics. The random forces on each domain are responsible for diffusing the system and creating the chaotic motion, but in order to have dynein prefer one side of the microtubule on average (as is the case for real dynein) we need to impose some \textit{extra} structure on the protein.

	To do this we steer each domain towards an equilibrium configuration by a Hooke's law restorative torque of the following form
	\begin{equation}
		\Gamma_i = c_i\left(\phi_i-\phi_{i,\text{eq}}\right)
	\end{equation}
	where the $c_i$ are spring constants (a free parameter of the model) and the set $\{\phi_i\}$ are the angles measured between the $i-1$ and $i+1$ domains. Note that for binding domains this angle may be measured to the microtubule or not measured if the domain is unbound from the microtubule. The angles $\{\phi_{i, \text{eq}}\}$ are the preferred angles of the system as illustrated in figure \ref{fig:torque angles}

	\begin{figure}[hbt!]
		\centering
		\includegraphics[angle=-90, width=0.75\columnwidth]{torque_angles}
		\caption{Example of dynein in the onebound configuration with torque angles $\{\phi_i\}$ shown. \textbf{NOTE: double check w/ David on how we select the angles e.g. do we take min of the two possible angles? How do we define the equilibrium angles?}}
		\label{fig:torque angles}
	\end{figure}

	\subsection{Determining constants of the model}
	I will insert discussion here about how we determined rod lengths and equilibrium angles from microscopy images. \\

	\begin{figure}[hbt!]
		\centering
		\includegraphics[width=0.3\columnwidth]{schematic-2-superimposed}
		\caption{Our model superimposed over microscopy image of Dynein \textbf{NOTE: add source} for determining lengths and equilibrium angles} 
		\label{fig:superimpmosed}
	\end{figure}
	\subsection{Equations of Motion}
	This is where I will discuss how we calculate the equations of motion for our two states-- One bound and both bound as well as how we algorithmically determine the interaction forces from the torques.

		\subsubsection{The one bound state}
		The following figure shows the model in the one bound configuration.\\
		\begin{figure}[!hbt]
			\centering
			\includegraphics[width=0.6\columnwidth]{onebound}
			\caption{The dynein model in the one bound state } 
		\end{figure}

		\noindent Here $\ell_S, \ell_T$ are fixed stalk and tail lengths. We have chosen the angles $\{\theta_i\}$ as they are the most convenient and lead to a description of the position of each domain that requires only one angle and an adjacent position, that is
		\begin{align}
		x's \quad &\begin{cases}
		x_0 = x_\text{bound} \\
		x_1 = x_0 + \ell_S \cos\theta_0 \\
		x_2 = x_1 + \ell_T \cos\theta_1 \\
		x_3 = x_2 + \ell_T \cos(\pi-\theta_3) = x_2 - \ell_T\cos\theta_3 \\
		x_4 = x_3 + \ell_S \cos(\pi-\theta_4) = x_3 - \ell_S\cos\theta_4
		\end{cases} \\
		y's \quad &\begin{cases}
		y_0 = 0 \\
		y_1 = y_0 + \ell_S\sin\theta_0 \\
		y_2 = y_1 + \ell_T\sin\theta_1 \\
		y_3 = y_2 - \ell_T\sin\theta_3 \\
		y_4 = y_3 - \ell_S\sin\theta_4 
		\end{cases}
		\end{align}

		\noindent Because equation (2) relates the forces on each domain to its velocity we need the time derivative of the above equations. 
		\begin{equation}
		\dot{x}'s \quad \begin{cases}
		\dot{x}_0 = 0 \\
		\dot{x}_1 = \dot{x}_0-\ell_S\sin\theta_0 \dot{\theta}_0 \\
		\dot{x}_2 = \dot{x}_1-\ell_T\sin\theta_1 \dot{\theta}_1 \\
		\dot{x}_3 = \dot{x}_2+\ell_T\sin\theta_3 \dot{\theta}_3 \\
		\dot{x}_4 = \dot{x}_3+\ell_S\sin\theta_4 \dot{\theta}_4
		\end{cases} \\
		\end{equation}
		\begin{equation}
		\dot{y}'s \quad \begin{cases}
		\dot{y}_0 = 0 \\
		\dot{y}_1 = \dot{y}_0 + \ell_S\cos\theta_0\dot{\theta}_0 \\
		\dot{y}_2 = \dot{y}_1 + \ell_T\cos\theta_1\dot{\theta}_1 \\
		\dot{y}_3 = \dot{y}_2 - \ell_T\cos\theta_3\dot{\theta}_3 \\
		\dot{y}_4 = \dot{y}_3 - \ell_S\cos\theta_4\dot{\theta}_4
		\end{cases} 
		\end{equation}

		\subsubsection{The both bound state}
		\subsubsection{Calculating Forces and Torques}
			Here I will discuss in detail how we calculate the forces and show how this is done in code 
	\subsection{The Code}
		\subsubsection{The Random Force}
\section{Testing the model} 
	\subsection{Sanity Checks}
	This sub section will be for the various methods we used to ensure our model is behaving reasonably 
		\subsubsection{Verifying size of time step} 
		Here I will talk about how we found autocorrelations functions for each domain energy at various values of dt in order to verify our time step is as large as can be
		\subsubsection{Energy conservation and the Equipartion theorem} 
		talk about code to check that we satisfy energy 
	\subsection{Fitting to other studies}