### 2019-01-22 papers/waczak-thesis/thesis/thesis.tex

\documentclass[12pt]{report}
\usepackage{graphicx} % make it easy to load in images
\graphicspath{{./images/}} % where to find images
\usepackage[a4paper, margin=1in]{geometry} % change margins 
\usepackage{fancyhdr} % add a header / footer 
\pagestyle{fancy}

\usepackage{amsmath}
\usepackage{amssymb}  
\usepackage{amsthm}
\usepackage{ulem}

\usepackage{caption}  % for creating simple subfigures
\usepackage{subcaption}
\title{
	{Simulating the Brownian Dynamics of a Two-Dimensional Model for the Dynein Motor Protein}\vspace{1.5em}\\
	\includegraphics[width=0.5\columnwidth]{OSU_logo.png}\vspace{0.5em}\\
	\large{Oregon State University}
}
\author{John Waczak}
\date{\today}

\begin{document}
	\maketitle

	\chapter*{Abstract}
	test
	\chapter*{Acknowledgments}
	I want to thank ...

	\tableofcontents

	\chapter{Introduction}
	\input{./chapters/introduction}
	\chapter{Dynein}
	\input{./chapters/dynein}
	\chapter{Methods}
	\input{./chapters/methods}

\bibliography{references}
\bibliographystyle{acm}
\end{document}