### 2019-01-29 papers/waczak-thesis/thesis/chapters/dynein.tex
more thesis stuff

\section{Overview}
\section{The microtubule}
\section{Drunken Walking}
\section{Transporting Cargo}
\section{Experimental Stepping Statistics}