### 2019-03-20 papers/waczak-thesis/thesis/chapters/results.tex

\todo[inline]{TODO: I plan to add description and section headings over break}
\todo[inline]{TODO: add discussion of powerstroke vs winch models. Discuss simple modifications to simulation parameters to try and fit for this alternate mechanochemical model}

\begin{figure}[!hbt]
	\centering
	\includegraphics[width=0.75\columnwidth]{paper_displacement_vs_step_length}
	\caption{\textbf{Two dimensional histogram} comparing the total signed step length to
			the signed initial displacement between the binding domains. A positive value
			indicates that the front foot made the step (leading). A negative value
			indicates the rear foot (lagging) made the step. Simulations show a strong,
			negative correlation. A similar fit from an experimental paper is plotted for comparison.}
\end{figure}

\begin{figure}[!hbt]
	\centering
	\includegraphics[angle=-90, width=1\columnwidth]{leading_vs_lagging}
	\caption{Diagram illustrating a leading foot step (a) and a lagging foot step (b).}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{paper_trajectory_plot}
	\caption{}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{paper_model_behavior}
	\caption{}
\end{figure}
\vspace{10em}

\todo[inline]{TODO: combine 5.1 with 5.2 to explain leading vs lagging}
\todo[inline]{TODO: consider removing leading lagging axis labels in 5.1 if redundant with figure caption} 
\todo[inline]{NOTE: it looks like there may be an issue with the middle most cartoon in 5.3. The tail does not appear to be the correct size}
\todo[inline]{TODO decrease overall width of tail in cartoon dynein. Increase thickness of stalk} 
\todo[inline]{TODO: Check to make sure time axis in final panel of fig 5.4 is actually a log scale}