### 2019-03-20 papers/waczak-thesis/thesis/chapters/theory.tex

Now that I have established some facts about motor proteins as well as specifics about dynein's many odd behaviors, we may begin our discussion of how to construct and test a physical model for its motion. First, we note that dynein's large mass of roughly 2MDa \cite{johnson_structure_1983} as compared to similar motor proteins such as Kinesin (a few hundred kDa) \cite{liao_kinesin_1998} exaggerates the effect of interactions with the  aqueous environment of the cell. This results in a much more significant drag force as well as a random element from water molecule collisions that changes the overall dynamics of the system. The set of equations governing this motion is called Brownian dynamics and is what we use to simulate the trajectory of dynein along the microtubule. 
	\section{Brownian dynamics}
	To model dynein we treat its various domains (binding, motor, and tail) as spheres held together by massless rigid-rods. In general, the dynamics of a constant mass particle are governed by Newton's second law
	\begin{equation}
		\sum_i \mathbf{F}_i = \dot{\mathbf{p}} = m\ddot{\mathbf{x}}
	\end{equation}
	Thus, determining the dynamics of a system boils down to describing the various forces which act upon it and then solving the resulting differential equation for some set of initial conditions. For small particles in a solvent, there are three important forces to consider. The first is linear drag, which we write as 
	\begin{equation}
		F_\text{drag} = -\gamma\dot{\mathbf{x}}
	\end{equation}
	so that the drag force is proportional to the velocity by a factor $\gamma$. For a spherical particle of radius $a$ moving through a solvent with liquid viscosity $\eta$ this was derived by Stoke to be 
	\begin{equation}
		\gamma = 6\pi\eta a
	\end{equation}

	The second characteristic force of Brownian dynamics is a random force due to density fluctuations in the solvent which we denote $F_\text{rand}$. Finally, we might also consider how individual particles interact with each other through some interaction force. For us, this will be accomplished through the rigid rods that outline the structure of the protein by fixing the inter-domain distances via tension. Putting all of these forces together results in what is often called the \textit{Langevin Equation}.
	\begin{equation}
		 -\gamma\dot{\mathbf{x}}+\mathbf{F}_\text{rand} + \mathbf{F}_\text{interact} = m\ddot{\mathbf{x}} 
	\end{equation}

	If we assume further that $|\gamma\dot{\mathbf{x}}|\gg |m\ddot{\mathbf{x}}|$, or in other words, we take the limit of strong drag, then we arrive at the definition of Brownian Dynamics. This allows us to neglect the acceleration term resulting in the following differential equation
	\begin{equation}
		\dot{\mathbf{x}} = \frac{1}{\gamma}\left( \mathbf{F}_\text{interact} + \mathbf{F}_\text{rand} \right)
	\end{equation}
	which we must now solve for the equations of motion of our model.

	\section{A Simplified Model for Dynein}
	As mentioned, we model dynein as a system of spherical domains held together by massless rigid-rods with the whole system moving according to Brownian dynamics. The random forces on each domain are responsible for diffusing the system and creating the expected chaotic motion, but in order to have dynein prefer walking towards one side of the microtubule, as is the case for real dynein, we need to impose some \textit{extra} structure on the protein.

	\begin{figure}[hbt!]
		\centering
		\includegraphics[angle=-90, width=0.6\columnwidth]{torque_angles}
		\caption{Example of dynein in the one bound configuration with torque angles $\{\varphi_i\}$ shown.}
		\label{fig:torque angles}
	\end{figure}

	To do this we steer each domain towards an equilibrium configuration by a Hooke's law restorative torque with signed magnitude of the following form
	\begin{equation}
		\Gamma_i = -c_i\left(\varphi_i-\varphi_{i,\text{eq}}\right)
	\end{equation}
	Here the $c_i$ are spring constants (a free parameter of the model) and the set $\{\varphi_i\}$ are the angles measured between the $i-1$ and $i+1$ domains. Equally useful is the harmonic energy associated with each torque. These are given by
	\begin{equation}
		U_i = \frac{1}{2}c_i(\varphi_i-\varphi_{i,\text{eq}})^2
	\end{equation}

	Real dynein is composed of a pair of identical heavy chains. Therefore, our angle convention is chosen starting at the $0^{th}$ domain in figure \ref{fig:torque angles}. The same convention is then repeated for the second leg with the following correspondence 
	\begin{align}
		\varphi_{0,eq} &\longleftrightarrow \varphi_{4,eq} \\ 
		\varphi_{1,eq} &\longleftrightarrow \varphi_{3,eq}
	\end{align}
	where the angles $\{\varphi_{i, \text{eq}}\}$ are the preferred angles of the system as illustrated in figure \ref{fig:torque angles}. The torques on each domain are related to forces by 
	\begin{equation}
		\mathbf{\Gamma} = \mathbf{r}\times\mathbf{F}
	\end{equation}
	so that knowing the distance between each domains means that we have specified each of the $\mathbf{\Gamma}_i$ and $\mathbf{r}_i$. In order to dynamically move each domain we must invert equation 4.10 to determine $\mathbf{F}_i$, however there is no unique way to invert such a cross product. Therefore, we must specify an algorithm for determining the forces that result from each torque.

	\section{Calculating Forces and Torques}
	However we decide to choose the forces on each domain, we would like them to satisfy  conservation angular momentum as this condition serves to keep the energy of the structure in check (add some stuff here) \todo[inline]{TODO: double check for other conditions on torques with David}

	Ultimately, we must specify the forces for three cases: a motor domain, and a tail domain, and the unbound binding domain. Having decided upon an algorithm for determining these forces, we could also calculate the force on a bound binding domain (attached to the microtubule) in order to estimate the maximum possible strain before unbinding occurs. This may prove interesting for future work but currently we do not care to calculate these forces as a bound binding domain is static and therefore, we do not need to evolve it's position.

	\begin{figure}[!hbt]
		\centering
		\includegraphics[angle =-90, width=0.65\columnwidth]{torque_forces.pdf}
		\caption{Toy model for three arbitrary domains. The force applied to adjacent domains due to the torque on the $i^{th}$ domain is shown. Forces are chosen to be perpendicular the inter-domain separation and in the direction that would move $\varphi_i$ towards its equilibrium configuration.}
		\label{fig:torque_force}
	\end{figure}

	Figure \ref{fig:torque_force} illustrates how the force on the motor domain due to the torque on the binding domain is chosen. In general, equation (3.10) suggests we should pick the magnitude of the force on the $i^{th}$ domain due to the torque on the $j=i\pm1$ domains to be
	\begin{equation}
		F_{ij} = \Gamma_j / \ell_{ij}
	\end{equation}
	where $\ell_{ij}$ is the distance between the $i$ and $j$ domains. We then make the choice to apply the force perpendicular to the inter-domain separation and in the direction of the equilibrium angle. Making this choice makes it clear that angular momentum will be conserved for the case where $\ell_{i\;i-1}=\ell_{i\;i+1}$. Finally to simplify computational complexity, we transform these two forces into forces on the $i^{th}$ domain by instead applying an equal and opposite force to each $\mathbf{F}_{ij}$ on the $i^{th}$ domain.

	With this algorithm, we can loop through each domain and calculate the torque-related forces on each domain. For binding domains, there is only one adjacent domain to consider. 
	\todo[inline]{TODO: add another figure or update above to indicate the equal-and-opposite forces that are actually kept. It may look nice to turn force arrows into dotted arrow and then add the force on the $i^{th}$ domain with a solid line}

	We are now armed with a way to calculate the forces on each domain due to the restorative torque. The only force which we are unable to calculate directly is the force due to tension. We will deal with these forces by applying the method Lagrange multipliers as we solve for the equations of motion.

	\section{Equations of Motion}
	The dynamics of each domain are determined by equation 3.5 so that in order to simulate the complicated structure we must solve a system of coupled first order differential equations. In doing so, it is convenient to introduce two states to distinguish between dynein's stepping and when both of its binding domains are attached to the microtubule. We shall refer to these configurations hereafter as the one bound state and the both bound state with the idea being that they directly correspond to the pre and post-stroke states defined by the mechanochemical cycle of \cite{cianfrocco_mechanism_2015}.

	In the one bound state, all but one of the domains are free to diffuse while the last is attached to the microtubule. In the both bound state, both binding domains are attached. This slight difference in geometry between the two scenarios leads to two independent equations motion motion that require careful consideration when transitioning between states. The following sections detail the derivation of these equations of motion for each configuration.

		\subsection{The one bound state}
		Figure \ref{fig:one bound} illustrates dynein in the one bound state with new angles $\{\theta_i\}$ chosen for convenience of calculation and computation.\\
		\begin{figure}[!hbt]
			\centering
			\includegraphics[angle=-90,width=0.75\columnwidth]{onebound}
			\caption{The dynein model in the one bound state }
			\label{fig:one bound}  
		\end{figure}

		\noindent Here $\ell_S, \ell_T$ are fixed stalk and tail lengths. With some simple trigonometry, the positions of each domain are given by
		\begin{align}
		x's \quad &\begin{cases}
		x_0 = x_\text{bound} \\
		x_1 = x_0 + \ell_S \cos\theta_0 \\
		x_2 = x_1 + \ell_T \cos\theta_1 \\
		x_3 = x_2 + \ell_T \cos(\pi-\theta_3) = x_2 - \ell_T\cos\theta_3 \\
		x_4 = x_3 + \ell_S \cos(\pi-\theta_4) = x_3 - \ell_S\cos\theta_4
		\end{cases} &
		y's \quad &\begin{cases}
		y_0 = 0 \\
		y_1 = y_0 + \ell_S\sin\theta_0 \\
		y_2 = y_1 + \ell_T\sin\theta_1 \\
		y_3 = y_2 - \ell_T\sin\theta_3 \\
		y_4 = y_3 - \ell_S\sin\theta_4 
		\end{cases}
		\end{align}

		\noindent Because equation (3.5) relates the forces on each domain to its velocity we need the time derivative of the above equations. 
		\begin{align}
		\dot{x}'s \quad \begin{cases}
		\dot{x}_0 = 0 \\
		\dot{x}_1 = \dot{x}_0-\ell_S\sin\theta_0 \dot{\theta}_0 \\
		\dot{x}_2 = \dot{x}_1-\ell_T\sin\theta_1 \dot{\theta}_1 \\
		\dot{x}_3 = \dot{x}_2+\ell_T\sin\theta_3 \dot{\theta}_3 \\
		\dot{x}_4 = \dot{x}_3+\ell_S\sin\theta_4 \dot{\theta}_4
		\end{cases} &
		\dot{y}'s \quad \begin{cases}
		\dot{y}_0 = 0 \\
		\dot{y}_1 = \dot{y}_0 + \ell_S\cos\theta_0\dot{\theta}_0 \\
		\dot{y}_2 = \dot{y}_1 + \ell_T\cos\theta_1\dot{\theta}_1 \\
		\dot{y}_3 = \dot{y}_2 - \ell_T\cos\theta_3\dot{\theta}_3 \\
		\dot{y}_4 = \dot{y}_3 - \ell_S\cos\theta_4\dot{\theta}_4
		\end{cases} 
		\end{align}

		The Brownian motion equation (3.5) also defines the velocity on each domain allowing us to write
		\begin{align}
			\dot{x}_i &= \frac{1}{\gamma_i}\Big(F_{i,\Gamma_i}^x+ \lambda_{i,i-1}(x_i-x_{i-1})+\lambda_{i,i+1}(x_i-x_{i+1})-F_{i,\text{rand}}^x\Big)\\
			\dot{y}_i &= \frac{1}{\gamma_i}\Big(F_{i,\Gamma_i}^y+ \lambda_{i,i-1}(y_i-y_{i-1})+\lambda_{i,i+1}(y_i-y_{i+1})-F_{i,\text{rand}}^y\Big) \notag
		\end{align}
		where superscript $x$ and $y$ denote the force components.

		\todo[inline]{Double check this calculation with David to make sure I have accounted for the tension terms correctly. For some reason, Elliott uses x's in both the x and y equations... Is this a typo?}

		 At this point, comparing equations 3.13 with 3.14 reveals we have identified 8 nontrivial equations (x and y derivatives; ignoring bound binding domain 0). There are four unknown derivatives $\{\dot\theta_i\}$ and four unknown Lagrange multipliers $\{\lambda_{ij}\}$ which altogether form a linear system which we solve using Mathematica. For a detailed calculation see the appendices.

		 \todo[inline]{TODO: add appendices with full matrix equations}

		\subsection{The both bound state}

		\section{Determining constants of the model}
		The stalk and tail lengths $\ell_S, \ell_T$ as well as a set of equilibrium angles $\{\varphi_{i,\text{eq}}\}$ were determined by examining microscopy images of dynein in various configurations. 
		\begin{figure}[hbt!]
			\centering
			\includegraphics[width=0.3\columnwidth]{schematic-2-superimposed}
			\caption{Our model superimposed over microscopy image of Dynein \textbf{NOTE: add source} for determining lengths and equilibrium angles} 
			\label{fig:superimpmosed}
		\end{figure}

	\section{Binding and Unbinding}

	\todo[inline]{Add theoretical description of binding and unbinding probabilities and their relation to tension and torque on binding domain. Explain the use of our exponential factor and why it is reasonable. Discuss potential to add in exponential factor to binding rate.}