### 2019-03-20 papers/waczak-thesis/thesis/thesis.tex

\documentclass[12pt]{report}
\usepackage{graphicx} % make it easy to load in images
\graphicspath{{./images/},{../../../plots/}} % where to find images
\usepackage[a4paper, margin=1in]{geometry} % change margins 
\usepackage{fancyhdr} % add a header / footer 
\pagestyle{fancy}

\usepackage{amsmath}
\usepackage{amssymb}  
\usepackage{amsthm}
\usepackage{ulem}

\usepackage{caption}  % for creating simple subfigures
\usepackage{subcaption}

\usepackage{titlesec} % redefine headings to remove "Chapter X"

\usepackage[labelfont=bf, textfont=md]{caption}

\usepackage{setspace}

\usepackage{hyperref}

\usepackage[color=cyan]{todonotes}

\usepackage[labelfont=bf]{caption}

\title{
	{Simulating the Brownian Dynamics of a Two-dimensional Model for the Dynein Motor Protein}\vspace{1.5em}\\
	\includegraphics[width=0.5\columnwidth]{OSU_logo.png}\vspace{0.5em}\\
	\large{Oregon State University}
}
\author{John Waczak}
\date{\today}

\begin{document}
	\onehalfspacing

	\maketitle

	\chapter*{Abstract}
	test
	\chapter*{Acknowledgments}
	I want to thank ...

	\tableofcontents

	\chapter{Introduction}
	\input{./chapters/introduction}
	\chapter{Dynein}
	\input{./chapters/dynein}
	\chapter{Theory}
	\input{./chapters/theory}
	\chapter{Methods}
	\input{./chapters/methods}
	\chapter{Results}
	\input{./chapters/results}
\bibliography{references}
\bibliographystyle{prl_like.bst}
\end{document}