### 2019-04-16 papers/waczak-thesis/thesis/thesis.tex

\documentclass[12pt]{report}
\usepackage{graphicx} % make it easy to load in images
\graphicspath{{./images/},{../../../plots/}} % where to find images
\usepackage[a4paper, margin=1in]{geometry} % change margins 
\usepackage{fancyhdr} % add a header / footer 
\pagestyle{fancy}

\usepackage{amsmath}
\usepackage{amssymb}  
\usepackage{amsthm}
\usepackage{ulem}

\usepackage{caption}  % for creating simple subfigures
\usepackage{subcaption}

\usepackage{titlesec} % redefine headings to remove "Chapter X"

\usepackage[labelfont=bf, textfont=md]{caption}

\usepackage{setspace}

\usepackage{hyperref}

\usepackage[color=cyan]{todonotes}

\usepackage[labelfont=bf]{caption}

\title{
	{Simulating the Brownian Dynamics of a Two-dimensional Model for the Dynein Motor Protein}\vspace{1.5em}\\
	\includegraphics[width=0.5\columnwidth]{OSU_logo.png}\vspace{0.5em}\\
	\large{Oregon State University}
}
\author{John Waczak}
\date{\today}

\begin{document}
	\onehalfspacing

	\maketitle

	\chapter*{Abstract}
	\vspace{-3 em}word count: \textbf{345}\\

	Cells are complicated structures whose components must perform controlled motion in order for the cell to execute its functions. Diffusion alone is far to slow of an action to produce the observed motions of cellular structures. The solution to this problem comes in the form of a class of proteins known as molecular motors that carry cargo throughout the cell by hydrolyzing adenosine triphosphate. Among these motors, the dynein protein is recognized for its odd stepping behaviors. Sometimes it steps forwards. Other times it steps backwards-- even been observed to occasionally shuffle. This peculiar motion makes dynein an interesting object of study, but its status as one a select few minus-end directed motors makes understanding the mechanics of dynein’s motion vital to our understanding of the cell. When these motors misbehave, fundamental processes like axonal transport and muscle contraction fail.

	Current observational techniques for studying biomolecules include techniques for imaging such as cryo-electron microscopy and x-ray crystallography, and techniques for tracking a protein’s motion such as fluorescent tagging combined with fluorescence microscopy.  These methods have revealed much about the different conformational states achieved by dynein during a step and have established a base of statistics for step lengths, velocities, generated forces, etc… Despite these advancements, the ability to image the full molecule for an entire stepping cycle remains beyond the capabilities of the current experimental apparatus.

	To address this gap, we have developed a simplified, two-dimensional model for the dynein molecule. We treat the protein as a system of massive domains held together by rigid rods steered towards equilibrium configurations by spring like torques and time evolved using Brownian dynamics. With our current set of parameters, the model displays directed motion with step length and velocity distributions ( give statistics here ) in agreement with experiment. Simulations also suggest that there is no correlation between the conformation of dynein before and after a step. These results indicate that a more computationally efficient simulation can be achieved with a Monte Carlo simulation for the both bound state coupled with the full Brownian simulation for the one bound state.

	\chapter*{Acknowledgments}
	I want to thank ...

	\tableofcontents

	\chapter{Introduction}
	\input{./chapters/introduction}
	\chapter{Dynein}
	\input{./chapters/dynein}
	\chapter{Theory}
	\input{./chapters/theory}
	\chapter{Methods}
	\input{./chapters/methods}
	\chapter{Results}
	\input{./chapters/results}
\bibliography{references}
\bibliographystyle{prl_like.bst}
\end{document}