### 2021-01-07 papers/jin-thesis/Chapters/5Results.tex

\section{Optimized Parameters}
\textit{Talk about best-fit parameters that matched experimentalists data. Include table with all parameters.}

\section{Stepping Patterns}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{/../../plots/mc_plots/u_final_disp_probability_distribution_30.0_5.50e+09_1.00e+08_0.2_1.05_0.0_90.0_197.0_242.0.png}
	\caption[Final Displacement Probability Distribution]{\textbf{Final Displacement Probability Distribution}}
	\label{fig:final_disp}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{Figures/step_length_hist.png}
	\caption[Step Length Histogram]{\textbf{Step Length Histogram}}
	\label{fig:step_length}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{Figures/prob_lagging_vs_init_L.png}
	\caption[Probability of Lagging Step]{\textbf{Probability of Lagging Step}}
	\label{fig:step_length}
\end{figure}

\textit{Analysis on stepping patterns and what the parameters of our model indicate. Does dynein take a lot of fast steps?}

\section{Agree with Experiment?}
\textit{Whole point of parameter fitting was to agree with Yildiz and/or other experimental data. Did we do a good job?}