### 2021-01-17 papers/jin-thesis/Chapters/5Conclusion.tex
Added a lot to my thesis directory

\section{Findings}
\textit{What do our results indicate? Do we think our model correctly represents real dynein and its stepping patterns?}

\section{Further Work}
\textit{How can we improve our model?}