### 2021-04-19 papers/jin-thesis/main.tex

\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{./Figures/},{../../plots/}}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{fancyhdr}
\pagestyle{fancy}

\usepackage{float}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{titlesec}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage[labelfont=bf, textfont=md]{caption}

\usepackage[square,numbers]{natbib}
\usepackage{indentfirst}

\input{../../data/paper_params.tex}

\floatstyle{plaintop}
\newfloat{algorithm}{thb}{lop}
\floatname{algorithm}{Algorithm}
\newenvironment{alg}
{\hrulefill\begin{enumerate}}
{\end{enumerate}\hrulefill}

\newcommand\davidsays[1]{\textbf{\it [D: #1]}}

\title{
	\Huge{Modeling the mechanisms of dynein's step using Monte Carlo simulations and Brownian dynamics}\vspace{1em}\\
	\Large{Jin Kiatvongcharoen} \vspace{0.5em}\\[2in] 
%	\includegraphics[width=0.4\columnwidth]{OSU_logo.png}\vspace{1em}\\
	\normalsize{An undergraduate thesis advised by Dr. David Roundy}\\
	\normalsize{submitted to the Department of Physics, Oregon State University}\\%\vspace{2em}\\
	\normalsize{in partial fulfillment of the requirements for the degree of}\\
	\normalsize{B.S. in Physics}\vspace{1.5 em}\\
	\normalsize{Draft submitted \today}\\

}
\date{}

\begin{document}
    \maketitle
    \onehalfspacing

	\chapter*{Abstract}
	Molecular motor proteins are the working soldiers within a cell. More specifically, the motor protein dynein delivers vital information along the microtubule for cellular processes like axonal transport and cell division. A failed delivery can cause severe damage to the cell’s functionality and lead to neurological diseases. Despite dynein's rich history in cellular research, the mechanisms of its motility remains poorly understood and is spectulated to be a cause for defects. In contrary to dynein having completely stochastic stepping, experimentalist Ahmet Yildiz and colleagues observed a dependence of dynein's stepping behavior as a function of interhead separation. To investigate this, we developed a two-dimensional, particle-rod model of dynein and simulated its stepping mechanisms using Monte Carlo methods and Brownian dynamics. The simulation optimized stepping patterns by fitting parameters according to Yildiz's results. In order to replicate the correlation observed by Yildiz, the binding rate was increased to encourage fast steps. Our results indicate that dynein takes many short and immediate steps that are undetectable to experimentalists in order to achieve stepping dependence on the initial domain separation. 
%	
%	 However, dynein's walk is unpredictable in nature and can vary in length and direction.
%	
%	 cargo that allows the cell to 
%	Molecular motor proteins are responsible for the cell’s ability to move, divide, and spatially organize itself. A defect within any of the three motor proteins can cause severe damage to the cell’s functionality and lead to neurological diseases. In order to fully understand the cause of such possible diseases, the motor protein dynein and more specifically its structure has been extensively studied. However, there has been limited research focused on the motion of dynein. To investigate this, we will simulate dynein’s stepping mechanisms using Monte Carlo methods and Brownian dynamics. The dynein will be computationally replicated with a particle-rod model undergoing Brownian motion in a fluid, where its stepping will be repeatedly simulated to generate an ensemble of statistics, via the Monte Carlo. The program will optimize dynein’s stepping patterns by fitting parameters according to experimental results of dynein’s step length, time, and probability of stepping.

    \begin{spacing}{1.2}
	\tableofcontents
	\listoffigures
	\end{spacing}

	\chapter{Introduction}
	\input{./Chapters/1Introduction}
	\chapter{Theory}
	\input{./Chapters/2Theory}
	\chapter{Methods}
	\input{./Chapters/3Methods}
	\chapter{Results \& Discussion}
	\input{./Chapters/4Results}
	\chapter{Conclusion}
	\input{./Chapters/5Conclusion}

\newpage
\cite{Burgess2003} \cite{Cianfrocco2015mechanism} \cite{Dewitt2012} \cite{Capek2017, waczak2019drunken}
\cite{GoodsellArt} \cite{eschbach2011cytoplasmic}
\cite{rao2019molecular} \cite{manna2020mechanistic} \cite{desantis2017lis1} 
\cite{elshenawy2020lis1} \cite{ando2020small} \cite{kinoshita2018step} \cite{qiu2012dynein} \cite{muller1973dynamic} \cite{lim2007vorticity} \cite{fehr2008kinesin} \cite{trott2018mathematical}

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}