### 2021-04-20 papers/jin-thesis/main.tex

\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{./Figures/},{../../plots/}}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{fancyhdr}
\pagestyle{fancy}

\usepackage{float}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{titlesec}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage[labelfont=bf, textfont=md]{caption}

\usepackage[square,numbers]{natbib}
\usepackage{indentfirst}

\input{../../data/paper_params.tex}

\floatstyle{plaintop}
\newfloat{algorithm}{thb}{lop}
\floatname{algorithm}{Algorithm}
\newenvironment{alg}
{\hrulefill\begin{enumerate}}
{\end{enumerate}\hrulefill}

\newcommand\davidsays[1]{\textbf{\it [D: #1]}}

\title{
	\Huge{Modeling the mechanisms of dynein's step using Monte Carlo simulations and Brownian dynamics}\vspace{1em}\\
	\Large{Jin Kiatvongcharoen} \vspace{0.5em}\\[2in] 
%	\includegraphics[width=0.4\columnwidth]{OSU_logo.png}\vspace{1em}\\
	\normalsize{An undergraduate thesis advised by Dr. David Roundy}\\
	\normalsize{submitted to the Department of Physics, Oregon State University}\\%\vspace{2em}\\
	\normalsize{in partial fulfillment of the requirements for the degree of}\\
	\normalsize{B.S. in Physics}\vspace{1.5 em}\\
	\normalsize{Draft submitted \today}\\

}
\date{}

\begin{document}
    \maketitle
    \onehalfspacing

	\chapter*{Abstract}
	The motor protein dynein is responsible for cellular processes such as axonal transport and cell division by delivering vital information along the microtubule track. A failed delivery can cause severe damage to the cell’s functionality and lead to neurological diseases. Despite dynein's rich history in cellular research, the mechanisms of its motility remains poorly understood due to its large size, flexible structure, and stochastic stepping. Therefore, recent studies attempts to further understand dynein's stepping mechanism in hopes to prevent motility failure. Experimentalist Ahmet Yildiz and colleagues observed a dependence within dynein's stepping behavior varying as a function of interhead separation. To investigate this, we developed a two-dimensional, particle-rod model of dynein and simulated its stepping mechanisms using Monte Carlo methods and Brownian dynamics. We emulated the powerstroke by simulating rate limiting steps that transition the model dynein between a both-bound and one-bound state. The simulation optimized parameter sets by fitting probability distributions to Yildiz's observations. In order to replicate the correlation observed by Yildiz, the binding rate was increased to encourage fast steps. Our results indicate that dynein takes many short and immediate steps that are undetectable to experimentalists in order to achieve stepping dependence on the initial domain separation. 
%	
%	 However, dynein's walk is unpredictable in nature and can vary in length and direction.
%	
%	 cargo that allows the cell to 
%	Molecular motor proteins are responsible for the cell’s ability to move, divide, and spatially organize itself. A defect within any of the three motor proteins can cause severe damage to the cell’s functionality and lead to neurological diseases. In order to fully understand the cause of such possible diseases, the motor protein dynein and more specifically its structure has been extensively studied. However, there has been limited research focused on the motion of dynein. To investigate this, we will simulate dynein’s stepping mechanisms using Monte Carlo methods and Brownian dynamics. The dynein will be computationally replicated with a particle-rod model undergoing Brownian motion in a fluid, where its stepping will be repeatedly simulated to generate an ensemble of statistics, via the Monte Carlo. The program will optimize dynein’s stepping patterns by fitting parameters according to experimental results of dynein’s step length, time, and probability of stepping.

    \begin{spacing}{1.2}
	\tableofcontents
	\listoffigures
	\end{spacing}

	\chapter{Introduction}
	\input{./Chapters/1Introduction}
	\chapter{Theory}
	\input{./Chapters/2Theory}
	\chapter{Methods}
	\input{./Chapters/3Methods}
	\chapter{Results \& Discussion}
	\input{./Chapters/4Results}
	\chapter{Conclusion}
	\input{./Chapters/5Conclusion}

\newpage
\cite{Burgess2003} \cite{Cianfrocco2015mechanism} \cite{Dewitt2012} \cite{Capek2017, waczak2019drunken}
\cite{GoodsellArt} \cite{eschbach2011cytoplasmic}
\cite{rao2019molecular} \cite{manna2020mechanistic} \cite{desantis2017lis1} 
\cite{elshenawy2020lis1} \cite{ando2020small} \cite{kinoshita2018step} \cite{qiu2012dynein} \cite{muller1973dynamic} \cite{lim2007vorticity} \cite{fehr2008kinesin} \cite{trott2018mathematical}

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}